---
title: "ltaw_analysis_moment"
author: "Bernice Cheung"
date: "2023-08-21"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(psych)
library(ggplot2)
library(stringr)
library(knitr)
library(lme4)
library(lmerTest)
library(nFactors)
library(corrplot)
library(janitor)
library(kableExtra)
library(GPArotation)
library(here)
library(RColorBrewer)
library(broom)
library(sjPlot)
```

# prep{.tab}

## load dataset{.tabset}
```{r}
survey_df <- read.csv(here("outputs", "ltaw_df_std_110823.csv")) %>%
  filter(flag_rec_include == T)
moment_df <- read.csv(here("outputs", "ltaw_moments_std_170823.csv")) %>%
  filter(flag_invalid_moment == FALSE) %>%
  filter(flag_dup_moment == FALSE)
```

## functions

```{r}
value_between <- function(vec, start, end){
  
  result <- vec[(which(vec == start)):(which(vec == end))]
  
  return(result)
}
```

## define aesthetics
```{r}
palette <- brewer.pal(n=8, name="Set2")
palette_sentiment = c(brewer.pal(n = 8,name="Blues")[8], "orange")
palette_groups = c("No" = brewer.pal(n = 8,name="Blues")[8],
                  "Yes" = "darkorange",
                  "full" = palette[1])
plot_aes = theme_minimal() +
  theme(legend.position = "top",
        legend.text = element_text(size = 8),
        text = element_text(size = 12, family = "Futura Medium"),
        axis.text = element_text(color = "black"),
        axis.ticks.y = element_blank())
```


# Moment Engagement

## Overall Descriptive
```{r}
dens_words <- density(moment_df$number_of_words)
df_dens_words <- data.frame(x=dens_words$x, y=dens_words$y)
probs <- c(0.05, 0.25, 0.5, 0.75, 0.95)
quantiles <- quantile(moment_df$number_of_words, prob=probs)
df_dens_words$quant <- factor(findInterval(df_dens_words$x,quantiles))
ggplot(df_dens_words, aes(x,y)) + geom_line() + geom_ribbon(aes(ymin=0, ymax=y, fill=quant)) + theme_classic() + theme(legend.position="none") + scale_fill_discrete(name = "",
    labels = c("0" = "0-5%",
               "1" = "6-25%",
               "2" = "26-50%",
               "3" = "51-75%",
               "4" = "76-95%",
               "5" = "95-100%")
  ) + labs(x = "Number of words per prompt", y = "") + scale_x_continuous(breaks = seq(0, 1800, by = 100), limits=c(0,1800)) +
  scale_x_continuous(limits = c(0,1200)) + 
  geom_vline(aes(xintercept = quantiles[1]), color = "red", size = 0.6) +
  geom_vline(aes(xintercept = quantiles[2]), color = "orange", size = 0.6) +
  geom_vline(aes(xintercept = quantiles[3]), color = "green", size = 0.6) +
  geom_vline(aes(xintercept = quantiles[4]), color="turquoise", size=0.6) +
  geom_vline(aes(xintercept = quantiles[5]), color="blue", size=0.6) +
  annotate("text", x=quantiles[1]-22, y=.0035, label="5%", size=3) +
  annotate("text", x=quantiles[2]-22, y=.0035, label="25%", size=3) +
  annotate("text", x=quantiles[3]-22, y=.0035, label="50%", size=3) +
  annotate("text", x=quantiles[4]-22, y=.0035, label="75%", size=3) +
  annotate("text", x=quantiles[5]-22, y=.0035, label="95%", size=3) +
  annotate("text", x=quantiles[3], y=.0005, label=paste("Median:\n",median(moment_df$number_of_words), "words"), size=3) 
```

```{r}
## Word Count Descriptive
dens_words <- density(as.numeric(moment_df$duration)/60)
df_dens_words <- data.frame(x=dens_words$x, y=dens_words$y)
probs <- c(0.05, 0.25, 0.5, 0.75, 0.95)
quantiles <- quantile(as.numeric(moment_df$duration)/60, prob=probs)
df_dens_words$quant <- factor(findInterval(df_dens_words$x,quantiles))
ggplot(df_dens_words, aes(x,y)) + geom_line() + geom_ribbon(aes(ymin=0, ymax=y, fill=quant)) + theme_classic() + theme(legend.position="none") + scale_fill_discrete(name = "",
    labels = c("0" = "0-5%",
               "1" = "6-25%",
               "2" = "26-50%",
               "3" = "51-75%",
               "4" = "76-95%",
               "5" = "95-100%")
  ) + labs(x = "Number of words per prompt", y = "") +
  scale_x_continuous(limits = c(0,6), breaks = seq(0, 6, 1)) + 
  geom_vline(aes(xintercept = quantiles[1]), color = "red", size = 0.6) +
  geom_vline(aes(xintercept = quantiles[2]), color = "orange", size = 0.6) +
  geom_vline(aes(xintercept = quantiles[3]), color = "green", size = 0.6) +
  geom_vline(aes(xintercept = quantiles[4]), color="turquoise", size=0.6) +
  geom_vline(aes(xintercept = quantiles[5]), color="blue", size=0.6) +
  annotate("text", x=quantiles[1]-22, y=.0035, label="5%", size=3) +
  annotate("text", x=quantiles[2]-22, y=.0035, label="25%", size=3) +
  annotate("text", x=quantiles[3]-22, y=.0035, label="50%", size=3) +
  annotate("text", x=quantiles[4]-22, y=.0035, label="75%", size=3) +
  annotate("text", x=quantiles[5]-22, y=.0035, label="95%", size=3) +
  annotate("text", x=quantiles[3], y=.1, label=paste("Median:\n",round(median(as.numeric(moment_df$duration)/60),2), "minutes"), size=3) 

```

## By prompt category

```{r}
prompt_category_des <- moment_df %>%
  group_by(prompt_category) %>%
  summarize(number_of_words = median(number_of_words, na.rm = T), 
            words_per_minute = median(words_per_minute, na.rm = T))
```

### number_of_words
```{r}
moment_df %>%
  ggplot(aes(x = number_of_words, y = prompt_category, fill = prompt_category)) +
  ggridges::geom_density_ridges(color = NA, scale = 1) +
  scale_x_continuous(limits = c(0,600)) +
  geom_label(data = prompt_category_des, 
             aes(y = prompt_category, x = number_of_words, label = sprintf("Median = %s", number_of_words)), fill = "white") +
  labs(x = "Word Count", y = "Prompt Category") +
  scale_fill_manual(values = palette) +
  plot_aes +
  theme(legend.position = "none")
```

### words_per_minute
```{r}
moment_df %>%
  ggplot(aes(x = words_per_minute, y = prompt_category, fill = prompt_category)) +
  ggridges::geom_density_ridges(color = NA, scale = 1) +
  scale_x_continuous(limits = c(30,230)) +
  geom_label(data = prompt_category_des, 
             aes(y = prompt_category, x = words_per_minute, label = sprintf("Median = %s", round(words_per_minute,2))), fill = "white") +
  labs(x = "Word Count", y = "Prompt Category") +
  scale_fill_manual(values = palette) +
  plot_aes +
  theme(legend.position = "none")
```


## By prompt construct

```{r}
prompt_construct_des <- moment_df %>%
  group_by(prompt_construct) %>%
  summarize(number_of_words = median(number_of_words, na.rm = T), 
            words_per_minute = median(words_per_minute, na.rm = T))
```

### number_of_words
```{r}
moment_df %>%
  ggplot(aes(x = number_of_words, y = prompt_construct, fill = prompt_construct)) +
  ggridges::geom_density_ridges(color = NA, scale = 1) +
  #scale_x_continuous(limits = c(0,600)) +
  #geom_label(data = prompt_construct_des, 
   #          aes(y = prompt_construct, x = number_of_words, label = sprintf("Median = %s", number_of_words)), fill = "white") +
  labs(x = "Word Count", y = "Prompt Category") +
  #scale_fill_manual(values = palette) +
  plot_aes +
  theme(legend.position = "none")
```

### speed
```{r}
moment_df %>%
  ggplot(aes(x = words_per_minute, y = prompt_construct, fill = prompt_construct)) +
  ggridges::geom_density_ridges(color = NA, scale = 1) +
  #scale_x_continuous(limits = c(0,600)) +
  #geom_label(data = prompt_construct_des, 
   #          aes(y = prompt_construct, x = number_of_words, label = sprintf("Median = %s", number_of_words)), fill = "white") +
  labs(x = "Speed", y = "Prompt Category") +
  #scale_fill_manual(values = palette) +
  plot_aes +
  theme(legend.position = "none")
```

# Explore factors associate with prompt engagement

```{r}
moment_merge <- moment_df %>%
  left_join(select(survey_df, userid, swls_tot, sris_insight_tot, aniq_tot, sris_reflection_tot), by = "userid")
```


## Variance partition{.tabset}

### word count

66% of variance is between subject
```{r}
icc_wordcount <- lmer(number_of_words ~ 1 + (1 | userid), data = moment_df)

VarCorr(icc_wordcount) %>%
  as_tibble() %>%
  mutate(ICC=vcov/sum(vcov)) %>%
  select(grp, ICC)
```

### speed

77% of variance is between subject
```{r}
icc_speed <- lmer(words_per_minute ~ 1 + (1 | userid), data = moment_df)

VarCorr(icc_speed) %>%
  as_tibble() %>%
  mutate(ICC=vcov/sum(vcov)) %>%
  select(grp, ICC)
```

## wellbeing & moment engagement{.tabscale}

### word counts
```{r}
mlm_count_1 <- lmer(scale(number_of_words) ~ scale(swls_tot) + (1 | userid), data = moment_merge)
summary(mlm_count_1)
```

```{r}
mlm_count_2 <- lmer(scale(number_of_words) ~ prompt_category + (1 | userid), data = moment_merge)
summary(mlm_count_2)
```

```{r}
plot_model(mlm_count_2, type = "pred")
```


```{r}
mlm_count_3 <- lmer(scale(number_of_words) ~ prompt_construct + (1 | userid), data = moment_merge)
summary(mlm_count_3)
```

```{r}
plot_model(mlm_count_3, type = "pred")
```


### speed
```{r}
mlm_speed_1 <- lmer(scale(words_per_minute) ~ scale(swls_tot) + (1 | userid), data = moment_merge)
summary(mlm_speed_1)
```

```{r}
mlm_speed_2 <- lmer(scale(words_per_minute) ~ scale(swls_tot) + prompt_category + (1 | userid), data = moment_merge)
summary(mlm_speed_2)
```

```{r}
moment_merge_std <- moment_merge %>%
  mutate(words_per_minute_std = scale(words_per_minute), 
         swls_tot_std = scale(swls_tot), 
         prompt_category = as.factor(prompt_category))


mlm_speed_3 <- lmer(words_per_minute_std ~ swls_tot_std * prompt_category + (1 | userid), data = moment_merge_std)
summary(mlm_speed_3)
```

```{r}
plot_model(mlm_speed_3, type = "int")
```

