---
title: "LTAW_Moment cleaning"
author: "Kristina Dale"
date: "2023-08-02"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
options(scipen=999)
```

# Scope
1. Basic cleaning of moment data
    Summary of moment completion
    Moment "scoring"
2. cleaned moments - csv

# prep {.tabset}
## load packages
```{r}
if(!require('pacman')) {
	install.packages('pacman')
}

pacman::p_load(tidyverse, labelled, devtools, plyr, haven, expss, DT, qwraps2, gtsummary, remotes, readxl, retidytext, openxlsx, reactable, reactablefmtr, ggwordcloud, topicmodels, here, install = TRUE)
```

## define aesthetics
```{r}
palette = c("#772e25", "#c44536", "#ee9b00", "#197278", "#283d3b", "#9CC5A1", "#ADA7C9", "grey50")
palette_type = c("#c44536", "#ee9b00", "#197278")
palette_pilot = c("#c44536", "#197278")
palette_sentiment = c(palette[2], palette[4])
plot_aes = theme_minimal() +
  theme(legend.position = "top",
        legend.text = element_text(size = 8),
        text = element_text(size = 12, family = "Futura Medium"),
        axis.text = element_text(color = "black"),
        axis.ticks.y = element_blank())
```

# load & tidy dataset {.tabset}
*NOTE for iteration: in the future i think the process of getting the moment data, regardless of source will be a bit different, and this step may be replaced by a database query*
moments collected via CameraTag, see study detail page for details[link]. Raw data comes from AWS transcription service, script from Mike M.
```{r}
# load moment data
#moments <- read.csv("/Users/kristina/Documents/Surveys/NEW /Lets Talk About Wellbeing/Raw Data/_a-b14b7c20-5f43-013b-1cdd-02c7c2460d41_full.csv")
moments <- read.csv(here("inputs", "_a-b14b7c20-5f43-013b-1cdd-02c7c2460d41_full.csv"))

# load prolific data
#prolific <- read.csv("/Users/kristina/Downloads/prolific_export_639e73b38884ae6555e593ca.csv")
prolific <- read.csv(here("inputs", "prolific_export_639e73b38884ae6555e593ca.csv")) 
```

## tidy datasets 

```{r}
# tidy moment dataset
moments_tidy <- moments %>%
  separate(name, into = c("userid", "itemid"), sep = "_", extra = "merge") %>% #seperate the name column into userid & itemid by the first "_"
  filter(str_detect(userid, "^[0-9]")) %>% # exclude testing cases by filtering only userid that starts with numbers
  filter(str_detect(itemid, "^(prompt|Prompt)")) %>% # include rows that has itemid starting with prompt
  select(-c(description, ip)) #remove empty & identifiable variables

# tidy prolific dataset
prolific_tidy <- prolific %>%
  rename_all(~paste0(., "_prolific"))  # add suffix
```

## merge datasets


```{r}
# merge prolific data into the moment dataset
moments_merge <- moments_tidy %>%
  left_join(prolific_tidy, by = c("userid" = "Participant.id_prolific")) # merge the two datasets by userid, which is named as "Participant.id_prolific" in the prolific dataset

uni_moment_id <- length(unique(moments_merge$userid))
uni_prolific_id <- length(unique(prolific_tidy$Participant.id_prolific))
```
`r uni_moment_id` out of `r uni_prolific_id` participants on prolific were included in this merge


# data screening {.tabset}

## exclude observations with revoked consent


```{r}
moments_consent <- moments_merge %>%
  replace_na(list(Status_prolific = "TIMEOUT")) %>% # replace "TIMEDOUT"
  filter(!Status_prolific == "RETURNED") # exclude observation with returned(revoked) consent 

moments_remove <- moments_merge %>% filter(Status_prolific == "RETURNED")
uni_exclude_id <- length(unique(moments_remove$userid))

```

`r nrow(moments_remove)` observations from `r uni_exclude_id` users are exclude from the dataset due to revoked consent.

## screen for valid moment

moments are screened based on item type and duration. The flag_invalid_moment:  
is TRUE if either the duration is none or the item type is PromptTest
is FALSE for all other moments
```{r}
moments_valid <- moments_consent %>%
  mutate(flag_invalid_moment = case_when(duration == "none" | itemid == "PromptTest" ~ TRUE,
                                  TRUE ~ FALSE))

table(moments_valid$flag_invalid_moment, useNA = "ifany")
```

## screen for duplicates

moments are screened based on duplicated item type within each participants. The flag_dup_moment: 
is TRUE if there are more than one identical item type within a userid (all duplicated moments are marked)
is NA if the moments are invalid 
is FALSE for all other moments.  


*In this particular study, participants were randomly presented with a subset (1/5 Narrative, 5/15 Well-being) of a larger pool of possible prompt options + 1 Goal prompt*, so it is possible that even though they accessed the study at multiple points, they did not complete the same prompt(s) twice due to randomization. Because of this, we flag only duplicated itemid's as dupes within a given participant. We retain all duplicated data and flag each response.

```{r}
moments_dup <- moments_valid %>%
  group_by(userid) %>%
  mutate(flag_dup_moment = case_when(flag_invalid_moment == TRUE ~ NA, 
                                     is_duplicate = duplicated(itemid) | duplicated(itemid, fromLast = TRUE) ~ TRUE, 
                                     TRUE ~ FALSE)) %>%
  ungroup()

table(moments_dup$flag_dup_moment, useNA = "ifany")

# store the cleaned dataset
moments_clean <- moments_dup
```

# Summry {.tabset}

## by prompt 
```{r}
moments_clean %>%
  filter(flag_invalid_moment == FALSE) %>%
  group_by(itemid) %>%
  dplyr::summarize(total_valid_moments = n())
```

## by prompt category
*NOTE: theres probably an easier way to do this where we have some larger prompt library sheet where all of these categories/labels are pre-defined, then we can read that in here and merge by itemid -- but for now, its done manually using the study outline*
```{r}
moments_clean$prompt_category <- ifelse(grepl("WB", moments_clean$itemid), "Well-being", "") 
moments_clean$prompt_category <- ifelse(grepl("Narr", moments_clean$itemid), "Narrative", moments_clean$prompt_category) 
moments_clean$prompt_category <- ifelse(grepl("Goal", moments_clean$itemid), "Goals", moments_clean$prompt_category) 
moments_clean$prompt_category <- ifelse(grepl("Test", moments_clean$itemid), "Test", moments_clean$prompt_category) 

moments_clean %>%
  filter(flag_invalid_moment == FALSE) %>%
  group_by(prompt_category) %>%
  dplyr::summarize(total_valid_moments = n())

```

## by prompt construct
```{r}
moments_clean <- moments_clean %>% #define constructs from study outline
  mutate(
    prompt_construct = case_when(
      itemid == "PromptNarr_1" ~ "Problem-solving",
      itemid == "PromptNarr_2" ~ "Identity",
      itemid == "PromptNarr_3" ~ "Bitterness revival",
      itemid == "PromptNarr_4" ~ "Competence building redemption",
      itemid == "PromptNarr_5" ~ "Competence building agency",
      itemid == "PromptWB_1" ~ "Autonomy",
      itemid == "PromptWB_2" ~ "Clear thinking",
      itemid == "PromptWB_3" ~ "Competence",
      itemid == "PromptWB_4" ~ "Emotional stability",
      itemid == "PromptWB_5" ~ "Empathy",
      itemid == "PromptWB_6" ~ "Engagement",
      itemid == "PromptWB_7" ~ "Meaning",
      itemid == "PromptWB_8" ~ "Optimism",
      itemid == "PromptWB_9" ~ "Positive emotions",
      itemid == "PromptWB_10" ~ "Positive relationships",
      itemid == "PromptWB_11" ~ "Prosocial behavior",
      itemid == "PromptWB_12" ~ "Resilience",
      itemid == "PromptWB_13" ~ "Self-acceptance",
      itemid == "PromptWB_14" ~ "Self-esteem",
      itemid == "PromptWB_15" ~ "Vitality",
      itemid == "PromptGoal" ~ "Goals",
      TRUE ~ NA_character_
    ))

moments_clean %>%
  filter(flag_invalid_moment == FALSE) %>%
  group_by(prompt_construct) %>%
  dplyr::summarize(total_valid_moments = n())

```

# label moment data
add item text column 
```{r}
moments_clean <- moments_clean %>% 
  mutate(
    itemtext = case_when(
      itemid == "PromptNarr_1" ~ "Talk about a time in your past when you successfully coped with a challenge.",
      itemid == "PromptNarr_2" ~ "Talk about an event or experience in your past that was meaningful and helped shape how you became the person that you are.",
      itemid == "PromptNarr_3" ~ "Talk about a negative experience in your past when you were disappointed in yourself, in conflict with someone else, or for some other reason have regrets.",
      itemid == "PromptNarr_4" ~ "Talk about a time you failed and at least one way this failure changed you for the better.",
      itemid == "PromptNarr_5" ~ "Talk about a time you succeeded and describe the steps you took to make this success a reality.",
      itemid == "PromptWB_1" ~ "What is keeping you from pursuing the things you want to pursue in life?",
      itemid == "PromptWB_2" ~ "Describe a time when you were very focused on a task.",
      itemid == "PromptWB_3" ~ "Talk about whether you feel able to manage difficulties in your life.",
      itemid == "PromptWB_4" ~ "Describe an event that recently made you upset. What did that feel like for you?",
      itemid == "PromptWB_5" ~ "Describe a time when you talked with another person about their problems.",
      itemid == "PromptWB_6" ~ "Describe a recent activity or hobby that you enjoyed.",
      itemid == "PromptWB_7" ~ "To what extent do you feel like your life is moving in the right direction?",
      itemid == "PromptWB_8" ~ "Do you feel optimistic about the future? Talk about it.",
      itemid == "PromptWB_9" ~ "Talk about something that usually makes you happy.",
      itemid == "PromptWB_10" ~ "Is there a community, online or in person, that you feel you have become a part of?",
      itemid == "PromptWB_11" ~ "Talk about a time when you sacrificed for someone else.",
      itemid == "PromptWB_12" ~ "Describe a time you had a significant life difficulty, and how you reacted to it.",
      itemid == "PromptWB_13" ~ "How do you treat yourself when you make mistakes?",
      itemid == "PromptWB_14" ~ "How would you describe yourself as a person?",
      itemid == "PromptWB_15" ~ "Talk about a time when you felt truly happy and alive.",
      itemid == "PromptGoal" ~ "If your goal was a more satisfying life, what would be your first step towards that goal?",
      itemid == "PromptTest" ~ "Using the microphone widget below, please record the following: Hello, this is a test recording.",
      TRUE ~ NA_character_
    ))

#organize
#moments_cleaned <- moments_tidy[c(1,3,39,4:12,32,36:38,13:31,33:35)]
#moments_cleaned <- moments_tidy
moments_clean$source <- "Lets talk about wellbeing"
```

# prompt level dataset

```{r}
prompt_df <- moments_clean %>%
  select(itemid, prompt_category,prompt_construct, itemtext) %>%
  unique() %>%
  filter(!prompt_category == "Test") %>%
  arrange(prompt_category) %>%
  mutate(prompt_id = sprintf("prompt_%02d", row_number())) %>%
  mutate(ltaw = TRUE)
```

```{r}
#write.csv(moments_cleaned, "/Users/kristina/Documents/Surveys/NEW /Lets Talk About Wellbeing/LTAW_moments_cleaned.csv")

#write.csv(moments_clean, here("outputs", "ltaw_moments_cleaned.csv"), row.names = F)
#write.csv(prompt_df, here("outputs", "prompt_dict.csv"), row.names = F)
```
