---
title: "LTAW_analysis_surveydata"
author: "Bernice Cheung"
date: "8/21/2023"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(psych)
library(ggplot2)
library(stringr)
library(knitr)
library(lme4)
library(nFactors)
library(corrplot)
library(janitor)
library(kableExtra)
library(GPArotation)
library(here)
library(RColorBrewer)
library(broom)
library(sjPlot)
```

# prep{.tab}

## load dataset{.tabset}
```{r}
survey_df <- read.csv(here("outputs", "ltaw_df_std_110823.csv"))
```

## functions

```{r}
value_between <- function(vec, start, end){
  
  result <- vec[(which(vec == start)):(which(vec == end))]
  
  return(result)
}
```

## pre-sets

```{r pre-sets}
var_names <- colnames(survey_df)
# well-being

wellbeing_items <- c(value_between(var_names, "swls_01", "pwb_00"), value_between(var_names, "wbpro_autonomy", "wbpro_vitality"))

wellbeing_subscale <- c(value_between(var_names, "pwb_autonomy_tot", "pwb_selfaccept_tot"), "swls_tot", "perma_pos", "perma_neg", value_between(var_names, "wbpro_autonomy", "wbpro_positive_emotions"), value_between(var_names, "wbpro_positive_relationships", "wbpro_vitality"))

wellbeing_single <- "pwb_00"

pwb_subscale_var <- value_between(var_names, "pwb_autonomy_tot", "pwb_selfaccept_tot")
wbpro_subscale_var <- c(value_between(var_names, "wbpro_autonomy", "wbpro_positive_emotions"), value_between(var_names, "wbpro_positive_relationships", "wbpro_vitality"))

main_dv <- c("swls_tot", "sris_reflection_tot", "sris_insight_tot", value_between(var_names, "aniq_awareness_tot", "aniq_tot"))

# total observation
obs_n <- nrow(survey_df)

# demo
demo_var <- c("demo_gender", "demo_age",value_between(var_names, "demo_education", "demo_family"), "language_prolific", "employment_status_prolific", "demo_race", "student_status_prolific")
```

## define aesthetics
```{r}
palette <- brewer.pal(n=8, name="Set2")
palette_sentiment = c(brewer.pal(n = 8,name="Blues")[8], "orange")
palette_groups = c("No" = brewer.pal(n = 8,name="Blues")[8],
                  "Yes" = "darkorange",
                  "full" = palette[1])
plot_aes = theme_minimal() +
  theme(legend.position = "top",
        legend.text = element_text(size = 8),
        text = element_text(size = 12, family = "Futura Medium"),
        axis.text = element_text(color = "black"),
        axis.ticks.y = element_blank())
```


# Well-being descriptive 

## overall well-being descriptive {.tabset}

```{r}
swls_mean <- mean(survey_df$swls_tot, na.rm = T)
swls_mean_ns <- mean(survey_df[survey_df$student_status_prolific == "No",]$swls_tot, na.rm = T)
swls_mean_s <- mean(survey_df[survey_df$student_status_prolific == "Yes",]$swls_tot, na.rm = T)
group_aes <- data.frame("group" = c("non-student", "student"), 
                        "group_color" = c(brewer.pal(n = 8,name="Blues")[8], "darkorange"), 
                        "group_mean" = c(swls_mean_ns, swls_mean_s))
swls_label <- c("Extremely Dissatisfied", "Dissatisfied", "Slightly Dissatisfied", "Neutral", "Slightly Satisfied", "Satisfied", "Extremely Satisfied")
```


### All participants
```{r}
survey_df %>%
  mutate(student_status_prolific = case_when(student_status_prolific == "DATA_EXPIRED" ~ NA, TRUE ~ student_status_prolific)) %>%
  filter(!is.na(swls_tot)) %>%
  ggplot(aes(x=swls_tot, fill = "full")) +
    geom_density(alpha=0.9, trim = TRUE) +
    scale_fill_manual(values = palette_groups) +
    scale_x_continuous(limits = c(5,35), breaks = seq(5, 35, 5), labels = swls_label) +
    geom_vline(aes(xintercept=swls_mean, color="black"),
               linetype="dashed", show.legend = FALSE) + 
    labs(x = "Overall Life Satisfaction") + 
    plot_aes + 
    theme(legend.position = "none", axis.text.x = element_text(angle=60, vjust = 0.5))
```

### non-studenet vs. student
non-student: 1581; student: 445; NA: 524

```{r}
survey_df %>%
  mutate(student_status_prolific = case_when(student_status_prolific == "DATA_EXPIRED" ~ NA, TRUE ~ student_status_prolific)) %>%
  filter(!is.na(swls_tot)) %>%
  filter(!is.na(student_status_prolific)) %>%
  ggplot(aes(x=swls_tot, fill = student_status_prolific)) +
    geom_density(alpha=0.9, trim = TRUE) +
    scale_x_continuous(limits = c(5,35), breaks = seq(5, 35, 5), labels = swls_label) +
    geom_vline(data= group_aes, aes(xintercept=group_mean, color=group_color),
              linetype="dashed", show.legend = FALSE) + 
    scale_fill_manual(values = palette_groups) +
    labs(x = "Overall Life Satisfaction") + 
    plot_aes + 
    theme(axis.text.x = element_text(angle=60, vjust = 0.5))
```

## Well-being dimensions{.tabset}

```{r}
wellbeing_subscale_mean <- survey_df %>%
  select(all_of(wellbeing_subscale)) %>%
  gather(dimension, ratings) %>%
  group_by(dimension) %>%
  summarize(dimension_mean = round(mean(ratings, na.rm = T), 2))
```


### Emotion
```{r}
survey_df %>%
  #filter(student_status_prolific == "No") %>%
  select(perma_neg,perma_pos) %>%
  gather(dimension, ratings) %>%
  ggplot(aes(x = ratings, y = dimension, fill = dimension)) +
  ggridges::geom_density_ridges(color = NA, scale = 1) +
  scale_x_continuous(limits = c(0,4), breaks = seq(0, 4, 1)) +
  geom_label(data = filter(wellbeing_subscale_mean, dimension %in% c("perma_neg", "perma_pos")), 
             aes(y = dimension, x = dimension_mean, label = sprintf("Mean = %s", dimension_mean)), fill = "white") +
  labs(x = "Ratings") +
  scale_fill_manual(values = palette_sentiment) +
  plot_aes +
  theme(legend.position = "none")
```

### PWB
```{r}
survey_df %>%
  #filter(student_status_prolific == "No") %>%
  select(all_of(pwb_subscale_var)) %>%
  gather(dimension, ratings) %>%
  ggplot(aes(x = ratings, y = dimension, fill = dimension)) +
  ggridges::geom_density_ridges(color = NA, scale = 1) +
  scale_x_continuous(limits = c(3,21), breaks = seq(3, 21, 3)) +
  geom_label(data = filter(wellbeing_subscale_mean, str_detect(dimension, "pwb_")), 
             aes(y = dimension, x = dimension_mean, label = sprintf("Mean = %s", dimension_mean)), fill = "white") +
  labs(x = "Ratings") +
  scale_fill_manual(values = palette) +
  plot_aes +
  theme(legend.position = "none")
```

### wbpro
```{r}
survey_df %>%
  #filter(student_status_prolific == "No") %>%
  select(all_of(wbpro_subscale_var)) %>%
  gather(dimension, ratings) %>%
  ggplot(aes(x = ratings, y = dimension, fill = dimension)) +
  ggridges::geom_density_ridges(color = NA, scale = 1) +
  scale_x_continuous(limits = c(1,9), breaks = seq(1, 9, 1)) +
  geom_label(data = filter(wellbeing_subscale_mean, str_detect(dimension, "wbpro_")), 
             aes(y = dimension, x = dimension_mean, label = sprintf("Mean = %s", dimension_mean)), fill = "white") +
  labs(x = "Ratings") +
  #scale_fill_manual(values = palette) +
  plot_aes +
  theme(legend.position = "none")
```

# Well-being Assessment

## Correlations between single item measures and well-being subscale/overall scores

```{r}
single_item_df <- survey_df %>% select(all_of(wellbeing_subscale), wellbeing_single)
single_cor_df <- cor(single_item_df, use = "pairwise", method = "spearman") %>%
  as.data.frame() %>%
  mutate(subscale = rownames(.)) %>%
  filter(subscale != "pwb_00") %>%
  arrange(desc(pwb_00)) 

single_cor_df$subscale <- reorder(rownames(single_cor_df), single_cor_df$pwb_00)

single_cor_df %>%
  ggplot(aes(x=subscale, y=pwb_00)) + 
  geom_col(fill = "darkorange") +
  geom_hline(yintercept=0, linetype="dashed") +
  labs(title="Correlation between 1-item wellbeing measures and other variables", y="Correlation coefficient") +
  coord_flip()  
```

## distribution comparison
```{r}
survey_df %>%
  mutate(swls_tot_std = scale(swls_tot), 
         pwb_00_std = scale(pwb_00)) %>%
  select(swls_tot_std, pwb_00_std) %>%
  gather(measure, scores) %>%
  ggplot(aes(x=scores, fill = measure)) +
    geom_density(alpha=0.9, trim = TRUE) +
    #scale_x_continuous(limits = c(5,35), breaks = seq(5, 35, 5), labels = swls_label) +
    scale_fill_manual(values = palette_sentiment) +
    labs(x = "Standardized score") + 
    plot_aes
```


## factor analysis

```{r}
# generate a correlational matrix
wellbeing_subscale_df <- survey_df %>%
  select(all_of(wellbeing_subscale)) %>%
  select(-swls_tot)

wellbeing_subscale_m <- cor(wellbeing_subscale_df, use = "pairwise")

# use Very Simple Structure criterion
res_vss <- psych :: nfactors(wellbeing_subscale_m, n = 10, rotate = "promax", diagonal = FALSE, fm = "minres", 
n.obs=obs_n,title="Very Simple Structure",use="pairwise",cor="cor")

# select useful parameters and organize them into a table
cbind(1:10, res_vss$map) %>%
  as_tibble() %>%
  rename(., factor = V1, map = V2) %>%
  cbind(., res_vss$vss.stats) %>%
  select(factor, map, fit, complex, eChisq, SRMR, eCRMS, eBIC, eRMS) %>%
  kable(format = "html", escape = F, caption = "VSS output -mTurk") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F,position = "center",fixed_thead = T)

# Use the Scree plot to identify the number of factors have Eigenvalues >1 and the output from the Parallel analysis

ev <- eigen(wellbeing_subscale_m)
ap <- parallel(subject=nrow(wellbeing_subscale_df),var=ncol(wellbeing_subscale_df),
  rep=100,cent=.05)
nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
plotnScree(nS)
```

### Extract 4 factors

```{r}
# extract 4 factors
fa_new_4 <-fa(r=wellbeing_subscale_m, nfactors=4,n.obs = obs_n, rotate="promax", SMC=FALSE, fm="minres")
```

```{r}
## Factor loadings
fa.diagram(fa_new_4)
```

The factors remain the same but some variables have shifted, such as intrinsic motivation
```{r}
# visualization
loadings <- fa.sort(fa_new_4)$loadings
loadings <- as.data.frame(unclass(loadings))
colnames(loadings) <- c("Positivity","Stability", "Autonomy", "Prosocial")
loadings$Items <- rownames(loadings)
loadings.m <- loadings %>% gather(-Items, key = "Factor", value = "Loading")
colOrder <- c("Positivity","Stability", "Autonomy", "Prosocial")
rowOrder <- rev(rownames(loadings))
loadings.m<- arrange(mutate(loadings.m,Items=factor(Items,leve=rowOrder)),Items)
loadings.m<- arrange(mutate(loadings.m,Factor=factor(Factor,leve=colOrder)),Factor)

ggplot(loadings.m, aes(Items, abs(Loading), fill=Loading)) + 
  facet_wrap(~ Factor, nrow=1) + #place the factors in separate facets
  geom_bar(stat="identity") + #make the bars
  coord_flip() + #flip the axes so the test names can be horizontal  
  #define the fill color gradient: blue=positive, red=negative
  scale_fill_gradient2(name = "Loading", 
                       high = "orange", mid = "white", low = "midnightblue", 
                       midpoint=0, guide="colourbar") +
  ylab("Loading Strength") + #improve y-axis label + 
  ggtitle("The Four Factors of current well-being assessment") +
  geom_hline(yintercept = 0.3, color = "red", linetype="dotted") +
  theme_bw(base_size=10)
```

# Demographic & well-being{.tabset}

1601/2878 complete cases
```{r}
demo_lm_df <- survey_df %>%
  select(all_of(demo_var), swls_tot) %>%
  mutate(across(demo_gender:student_status_prolific, ~ case_when(
    . == "DATA_EXPIRED" ~ NA_character_,
    . == "I prefer not to answer" ~ NA_character_,
    TRUE ~ as.character(.)
  ))) %>%
  mutate(demo_race_bi = case_when(demo_race != "White/Caucasian" & !is.na(demo_race) ~ "Non-White", 
                                  TRUE ~ demo_race)) %>%
  filter(complete.cases(.))
```


## ANOVA{.tabset}

### Individual demographic variable
```{r}
demo_lm_df %>%
  gather(demo_var, value, -swls_tot) %>%
  nest(-demo_var) %>%
  mutate(fit = map(data, ~ aov(swls_tot ~ value, data = .x)), 
         sum = map(fit, summary), 
         tidied = map(fit, tidy)) %>%
  unnest(tidied) %>%
  filter(term == "value") %>%
  select(demo_var, statistic, p.value) %>%
  mutate_if(is.numeric, round, 3) %>%
  arrange(desc(statistic)) %>%
  kable(format = "html", escape = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F,position = "center")
```

### Interaction with gender
```{r}
demo_lm_df %>%
  gather(demo_var, value, -swls_tot, -demo_gender) %>%
  nest(-demo_var) %>%
  mutate(fit = map(data, ~ aov(swls_tot ~ value *demo_gender, data = .x)), 
         sum = map(fit, summary), 
         tidied = map(fit, tidy)) %>%
  unnest(tidied) %>%
  filter(str_detect(term, "value:")) %>%
  select(demo_var, statistic, p.value) %>%
  mutate_if(is.numeric, round, 3) %>%
  arrange(desc(statistic)) %>%
  kable(format = "html", escape = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F,position = "center")
```

### Interaction with age
```{r}
demo_lm_df %>%
  gather(demo_var, value, -swls_tot, -demo_age) %>%
  nest(-demo_var) %>%
  mutate(fit = map(data, ~ aov(swls_tot ~ value *demo_age, data = .x)), 
         sum = map(fit, summary), 
         tidied = map(fit, tidy)) %>%
  unnest(tidied) %>%
  filter(str_detect(term, "value:")) %>%
  select(demo_var, statistic, p.value) %>%
  mutate_if(is.numeric, round, 3) %>%
  arrange(desc(statistic)) %>%
  kable(format = "html", escape = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F,position = "center")
```
## Interaction with race
```{r}
demo_lm_df %>%
  gather(demo_var, value, -swls_tot, -demo_race_bi) %>%
  nest(-demo_var) %>%
  mutate(fit = map(data, ~ aov(swls_tot ~ value *demo_race_bi, data = .x)), 
         sum = map(fit, summary), 
         tidied = map(fit, tidy)) %>%
  unnest(tidied) %>%
  filter(str_detect(term, "value:")) %>%
  select(demo_var, statistic, p.value) %>%
  mutate_if(is.numeric, round, 3) %>%
  arrange(desc(statistic))
```

## Visualization{.tabset}

### Income & gender
```{r}
demo_lm_df %>%
  ggplot(
  aes(x = demo_marital, y = swls_tot, fill = demo_gender)
) +
  geom_bar(position = 'dodge', stat = 'summary', fun = 'mean', alpha=0.7, colour=1) +
  #scale_fill_manual(values = palette) +
    labs(x = "Marital Status", y = "Overall Life Satisfaction") + 
    plot_aes

```

```{r}
demo_lm_df %>%
  mutate(demo_income = factor(demo_lm_df$demo_income, levels = c("Less than $25,000", "$25,000 - $50,000", "$51,000 - $100,000", "$101,000 - $200,000", "More than $200,000"))) %>%
  ggplot(
  aes(x = demo_income, y = swls_tot, fill = demo_income)
) +
  geom_bar(position = 'dodge', stat = 'summary', fun = 'mean', alpha=0.7, colour=1) +
  scale_fill_manual(values = palette) +
    labs(x = "Income", y = "Overall Life Satisfaction") + 
    plot_aes +
  theme(legend.position = "none")
  
```

```{r}
demo_lm_df %>%
  #mutate(demo_income = factor(demo_lm_df$demo_income, levels = c("Less than $25,000", "$25,000 - $50,000", "$51,000 - $100,000", "$101,000 - $200,000", "More than $200,000"))) %>%
  ggplot(
  aes(x = demo_race, y = swls_tot, fill = demo_race)
) +
  geom_bar(position = 'dodge', stat = 'summary', fun = 'mean', alpha=0.7, colour=1) +
  scale_fill_manual(values = palette) +
    labs(x = "Race", y = "Overall Life Satisfaction") + 
    plot_aes +
  theme(legend.position = "none", axis.text.x = element_text(angle=60, vjust = 0.5))
```

```{r}
demo_lm_df %>%
  #mutate(demo_income = factor(demo_lm_df$demo_income, levels = c("Less than $25,000", "$25,000 - $50,000", "$51,000 - $100,000", "$101,000 - $200,000", "More than $200,000"))) %>%
  ggplot(
  aes(x = demo_age, y = swls_tot, fill = demo_age)
) +
  geom_bar(position = 'dodge', stat = 'summary', fun = 'mean', alpha=0.7, colour=1) +
  scale_fill_manual(values = palette) +
    labs(x = "Age", y = "Overall Life Satisfaction") + 
    plot_aes +
  theme(legend.position = "none", axis.text.x = element_text(angle=60, vjust = 0.5))
```

# helpful resources{.tabset}

## Full sample
```{r}
survey_df %>%
  select(contains("resourcehelp_")) %>%
  select(-"resourcehelp_09_text") %>%
  gather(resourcehelp) %>%
  filter(!is.na(value)) %>%
  mutate(value = reorder(value, value, function(x) -length(x))) %>%
  ggplot(aes(x = value)) +
  geom_bar(fill = "darkorange") + 
  scale_fill_manual(values = palette) + 
  labs(x = "Helpful Resources to improve life satisfaction") +
  plot_aes +
  theme(axis.text.x = element_text(angle=60, vjust = 0.5))
  
```

## students
```{r}
survey_df %>%
  select(contains("resourcehelp_"), student_status_prolific) %>%
  select(-"resourcehelp_09_text") %>%
  filter(student_status_prolific == "Yes") %>%
  select(-student_status_prolific) %>%
  gather(resourcehelp) %>%
  filter(!is.na(value)) %>%
  mutate(value = reorder(value, value, function(x) -length(x))) %>%
  ggplot(aes(x = value)) +
  geom_bar(fill = "darkorange") + 
  scale_fill_manual(values = palette) + 
  plot_aes +
  theme(axis.text.x = element_text(angle=60, vjust = 0.5))
```



# helpful content{.tabset}

## Full Sample
```{r}
survey_df %>%
  select(contains("contenthelp_")) %>%
  select(-"contenthelp_12_text") %>%
  gather(contenthelp) %>%
  filter(!is.na(value)) %>%
  mutate(value = reorder(value, value, function(x) -length(x))) %>%
  ggplot(aes(x = value)) +
  geom_bar(fill = "darkorange") + 
  scale_fill_manual(values = palette) + 
  plot_aes +
  theme(axis.text.x = element_text(angle=60, vjust = 0.5))
```
## students
```{r}
survey_df %>%
  select(contains("contenthelp_"), student_status_prolific) %>%
  select(-"contenthelp_12_text") %>%
  filter(student_status_prolific == "Yes") %>%
  select(-student_status_prolific) %>%
  gather(contenthelp) %>%
  filter(!is.na(value)) %>%
  mutate(value = reorder(value, value, function(x) -length(x))) %>%
  ggplot(aes(x = value)) +
  geom_bar(fill = "darkorange") + 
  scale_fill_manual(values = palette) + 
  plot_aes +
  theme(axis.text.x = element_text(angle=60, vjust = 0.5))

```

# Associations between self-awareness, narrative identity & wellbeing

```{r}
dv_cdf <- survey_df %>% select(all_of(main_dv))
cor(dv_cdf, use = "pairwise", method = "spearman") %>%
  as.data.frame() %>%
  mutate(subscale = rownames(.)) %>%
  filter(subscale != "swls_tot") %>%
  ggplot(aes(x=subscale, y=swls_tot)) + 
  geom_col() +
  geom_hline(yintercept=0, linetype="dashed") +
  labs(title="Correlation between X and other variables", y="Correlation coefficient") +
  coord_flip()  
```

```{r}
m1 <- lm(swls_tot ~ sris_reflection_tot + sris_insight_tot + aniq_tot, data = survey_df)
summary(m1)
```

