---
title: "ltaw_standardize"
author: "`r Sys.getenv('USER')`"
date: '`r Sys.Date()`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, message = FALSE)
```

Standardize Let’s Talk About Well-being dataset based on the variable name and demographic dictionary. 

# Prepping {.tabset}

## load libraries
```{r libraries}
library(tidyverse)
library(here)
library(psych)
library(janitor)
library(broom)
library(gt)
library(gtsummary)
library(dplyr)
```

## load data
```{r}
# load cleaned datasets
ltaw_df <- read.csv(here("outputs", "LTAW_data_scored.csv"))

# load item variable dictionary
#varname_dic <- read.csv(here("outputs", "varname_dic_030823.csv"))

# load demographic level dictionary
#demo_dic <- read.csv(here("outputs", "demo_level_dic_020823.csv"))
```

## org_varname function
A function to organize variable names
```{r}
org_varname <- function(dataframe){
  scale_df <- data.frame("var_name" = colnames(dataframe), 
                            "var_class" = sapply(dataframe, class)) %>%
  separate(var_name, into = c("scale_prefix", "item_suffix"), sep = "_", remove = FALSE, extra = "merge") %>%
  mutate(num_item = grepl("\\d", item_suffix))
  
  scale_df <- scale_df %>%
  mutate(item_type = case_when(
    str_detect(scale_prefix, "(?i)Consent") ~ "metadata",
    str_detect(scale_prefix, "(?i)recording") ~ "metadata",
    str_detect(scale_prefix, "(?i)demo") ~ "demo",
    str_detect(scale_prefix, "(?i)flag") ~ "data_screen_flags",
    str_detect(scale_prefix, "(?i)max") ~ "data_screen_flags",
    str_detect(scale_prefix, "(?i)Attention") ~ "attention_check",
    str_detect(scale_prefix, "(?i)Prompt") ~ "prompt",
    str_detect(item_suffix, "(?i)_tot") ~ "subscale",
    str_detect(item_suffix, "(?i)_R") ~ "recoded_item",
    str_detect(item_suffix, "(?i)prolific") ~ "prolific_metadata", 
    str_detect(scale_prefix, "(?i)flag") ~ "data_screen_flags",
    num_item ~ "item",
    !is.na(item_suffix) & !num_item & var_class != "character" ~ "subscale",
    scale_prefix == "outro" ~ "feedback", 
    TRUE ~ "metadata"
  ))
  
  return(select(scale_df, -num_item))
}
```

## compare_varname function

A function to compare the variable names from the dataset against the dictionary
```{r}
compare_varname <- function(varname_dic, df){
  
  
  # add row index to the project dataframe
  df_compare <- df %>%
    mutate(row_idx = row_number()) %>%
    select(var_name, var_class, row_idx)
  
  # merge the project dataframe with the variable name dictionary
  matched_df <- varname_dic %>%
    left_join(df_compare, by = c("common_name" = "var_name"), suffix =c("_dic", "_df"))
  
  # identify unmatched rows from the project dataframe
  unmatched_df <- anti_join(df_compare, varname_dic, by = c("var_name" = "common_name"))
  
  return(list(matched_df = matched_df, unmatched_df = unmatched_df))
  
}
```

## rename_match function

A function to rename variable names based on the naming convention

```{r}
rename_var_std <- function(df) {
  df %>%
    #clean_names() %>% # use janitor::clean_names for a more thorough cleanup
    rename_with(~ tolower(gsub(".", "_", .x, fixed = TRUE))) %>% # replace "." by "_" and change everything into lower case
    rename_with(~ str_replace(., "(\\d+)", function(x) sprintf("%02d", as.numeric(x))), .cols = everything()) #add padding 0 to all numeric index
}
```

## std_var_name function

A function to standardize variable names. It renames matched variables based on the dictionary and renames unmatched variables based on the naming convention
```{r}
std_var_name <- function(df, varname_dic){
  
  # organize variable name
  varname_df <- org_varname(df)
  
  # rename matched variable
  # compare the project variable against the variable name dictionary
  df_compare_results <- compare_varname(varname_dic, varname_df)
  # subset a dataset with comparable variables 
  matched_df <- df_compare_results$matched_df
  # rename the variable names based on the matched names from the dictionary
  colnames(df)[matched_df$row_idx] <- matched_df$std_varname
  
  # rename unmatched variable
  df_rn <- rename_var_std(df)
  
  # organize standardized variable name 
  varname_df_std <- org_varname(df_rn)  %>% # create a summary table
    mutate(ori_name = varname_df$var_name) %>%
    mutate(match_dic = row_number() %in% matched_df$row_idx)
  
  std_results <- list(df_std = df_rn, varname_sum_std = varname_df_std)
  return(std_results)
}
```

## value_between

```{r}
value_between <- function(vec, start, end){
  
  result <- vec[(which(vec == start)):(which(vec == end))]
  
  return(result)
}
```

## as_factor_same

```{r}
as_factor_same <- function(df, var_names){
  df_f <- df %>%
    mutate(across(var_names, ~as.factor(.)))
  
  return(df_f)
}
```

## as_factor_recode
```{r}
as_factor_recode <- function(df, var_names, vec_recode_pattern){
  df_f <- df %>%
    mutate(across(var_names, ~as.factor(.))) %>%
    mutate(across(var_names, ~recode_factor(., !!!vec_recode_pattern)))
  
  return(df_f)
}
```


## demo_summary function

A function to summarize demographic variables and levels
```{r}
demo_summary <- function(dataframe){
  
  # convert all demo variables into factors
  demo_df <- dataframe %>%
    select(contains("Demo")) %>%
    mutate_all(as.factor)
  
  # extract levels and their total number
  demo_levels <- sapply(demo_df, levels)
  demo_levels_n <- sapply(demo_levels, length)
  
  # transform the levels into a dataframe
  demo_levels_df <- map_dfr(names(demo_levels), ~ data.frame(demo_var = .x, levels = unlist(demo_levels[[.x]]))) %>%
    nest(demo_levels = c(levels), .by = demo_var, .key = "demo_level_name")
  
  # combine
  demo_sum <- data.frame("demo_var" = colnames(demo_df), 
                                 "demo_levels_n" = demo_levels_n) %>%
    left_join(demo_levels_df, by = "demo_var") %>%
    unnest(demo_levels)
  
  return (demo_sum)
}
```

## compare_demo function

A function to compare demographic levels against the dictionary
```{r}
compare_demo <- function(demo_dic, df){
  
  # organize demographic levels
  demo_level_df <- demo_summary(df)

  # compare demographic variables and number of levels
  demo_dic_sum <- demo_dic %>%
    nest(levels) %>%
    rename(demo_levels = data)
  
  demo_df_sum <- demo_level_df %>%
    nest(levels) %>%
    rename(demo_levels = data)
  
  demo_sum_compare <- demo_dic_sum %>%
    full_join(demo_df_sum, by = c("common_name" = "demo_var"), suffix = c("_dic", "_df"))
  
  # check demographic variables that are not included in the dataset 
  missing_demo_var <- anti_join(demo_dic_sum, demo_df_sum, by = c("common_name" = "demo_var"))
  
  if(nrow(missing_demo_var) != 0) {
    missing_demo <- sprintf("Demographic variables that are not included in the dataset: %s \n", missing_demo_var$common_name)
    cat(missing_demo)
  } else {
    cat("All demographic variables from the dictionary are included in the dataset \n")
  } 
  
  # check demographic variables that are not included in the dictionary
  additional_demo_var <- anti_join(demo_df_sum, demo_dic_sum, by = c("demo_var" = "common_name"))

  if(nrow(additional_demo_var) != 0) {
    missing_demo <- sprintf("Demographic variables that are not included in the dictionary: %s \n", additional_demo_var$demo_var)
    cat(missing_demo)
  } else {
    cat("No additional demographic varialbe is included in the dataset \n")
  } 
  
  
  # compare items & subscales
  demo_compare_eval <- demo_sum_compare %>% 
  mutate(
         same_demo_n = identical(demo_levels_n_dic, demo_levels_n_df), # check if the index of items are identical
         same_demo_level = identical(demo_levels_dic, demo_levels_df)
         ) # check if the numes of subscales are identical
  
  return(demo_compare_eval)
  
}
```

## std_demo_level
```{r}
std_demo_level <- function(df, demo_dic) {
  
  # standardize demographic levels based on naming conventions
  df_rl <- df %>%
    mutate_at(vars(starts_with("demo_")), ~replace(., . == "", NA)) %>% # replace empty strings with NA
    mutate_at(vars(starts_with("demo_")), ~str_to_lower(.)) # transform all categories into lower_case
  
  # compare demographic levels against dictionary
  demo_sum_compare <- compare_demo(demo_dic, df_rl)
  
  # generate a summary of demographic levels
  demo_level_sum <- demo_summary(df_rl)
  
  std_demo_l <- list(df_rl = df_rl, demo_level_sum = demo_level_sum)
  
  return(std_demo_l)
  
}
```

## factor_summary
```{r}
factor_summary <- function(df, varname_sum){
  
  # extract factor name
  factor_var_df <- varname_sum %>%
  filter(var_class == "factor") %>%
  filter(!str_detect(var_name, "demo_"))

  factor_varname <- factor_var_df$var_name
  
  # extract factor data
  factor_df <- df %>%
  select(one_of(factor_varname))

  # extract levels and their total number
  factor_levels <- sapply(factor_df, levels)
  factor_levels_n <- sapply(factor_levels, length)
  
   factor_levels_df <- map_dfr(names(factor_levels), ~ data.frame(var_name = .x, levels = unlist(factor_levels[[.x]]))) %>%
    nest(factor_levels = c(levels), .by = var_name, .key = "demo_level_name")
  
  # combine
  factor_sum <- data.frame("var_name" = factor_var_df$var_name, 
                                 "factor_levels_n" = factor_levels_n) %>%
    left_join(factor_levels_df, by = "var_name")
  
  return(factor_sum)
  
}

```

# A summery of orignal variable names

```{r}
ori_varname <- org_varname(ltaw_df)

ori_varname %>%
  gt(caption = "Let's Talk About Wellbeing Original Variables")
```

# change item class

## recode character to factor without changing levels
```{r}
# check all character variables
ltaw_char_df <- ori_varname %>%
  filter(var_class == "character")

ltaw_char_df %>%
  gt(caption = "All character variables")

# identify variables that need to be recoded as factors
chr_col_names <- ltaw_char_df$var_name 

f_varname <- c(value_between(chr_col_names, "Demo_gender", "ResourceHelp_9"),
               value_between(chr_col_names, "ContentHelp_1", "ContentHelp_12"), 
               value_between(chr_col_names, "Languishing_1", "Document_Tools_7"), 
               value_between(chr_col_names, "Document_Freq", "Document_Benefits_11"), 
               value_between(chr_col_names, "Sex_prolific", "SWLS_tot_f"))

# recode them as factors
ltaw_df_rc <- as_factor_same(ltaw_df, f_varname)
```

## recode integer to factor 
```{r}
# check all integer variables
ltaw_int_df <- ori_varname %>%
  filter(var_class == "integer")

ltaw_int_df %>%
  gt(caption = "All integer variables")

# recode integer into boolean or factor (1,2)
new_levels <- c("1" = TRUE, "2" = FALSE)
recode_varname <- grep("_confirm", colnames(ltaw_df), value = TRUE)
ltaw_df_rc <- as_factor_recode(ltaw_df_rc, recode_varname, new_levels)

# recode integer into boolean or factor (0,1)
new_levels <- c("1" = TRUE, "0" = FALSE)
ltaw_df_rc <- as_factor_recode(ltaw_df_rc, c("Recording_test", "content_recruit"), new_levels)

# check all variable class
org_varname(ltaw_df_rc) %>%gt()
```


# Standardize variable names

```{r}
# rename variable according to the naming convention
ltaw_df_rn <- rename_var_std(ltaw_df_rc)

# create a summary table
ltaw_rn_sum <- org_varname(ltaw_df_rn) %>%
  mutate(ori_name = names(ltaw_df_rc))

ltaw_rn_sum %>% gt()
```



```{r}
# # standardize variable names 
# ltaw_std_name_l <- std_var_name(ltaw_df_rc, varname_dic)
# 
# # extract renamed dataframe
# ltaw_df_rn <- ltaw_std_name_l$df_std
# 
# # extract a summary of renamed variables
# ltaw_rn_sum <- ltaw_std_name_l$varname_sum_std
# 
# ltaw_rn_sum %>%
#   gt(caption = "Let's Talk About Well-being Standardized Variable Name Summary")
```


# change item type
```{r}
# extract variables related to content recommendation and goal-setting
content_rec_varname_raw <- value_between(ltaw_rn_sum$var_name, "resourcehelp_01", "contenthelp_12_text")
content_rec_varname <- c(value_between(ltaw_rn_sum$var_name, "promptgoal_confirm", "promptgoal_time_click_count"), "resourcehelp_09_text", "contenthelp_12_text")

# extract variables related to value prop
value_prop_varname_raw <- value_between(ltaw_rn_sum$var_name, "languishing_01", "document_benefits_11")
value_prop_varname <- c(value_between(ltaw_rn_sum$var_name, "document_benefits_11_text", "storyinterest_desc"), "document_tools_06_text", "document_freq")

prolific_demo <- value_between(ltaw_rn_sum$var_name, "fluent_languages_prolific", "employment_status_prolific")

ltaw_rn_sum <- ltaw_rn_sum %>%
  mutate(item_type = case_when(var_name %in% content_rec_varname ~ "content_rec", 
                               var_name %in% content_rec_varname_raw ~ "content_rec_raw", 
                               var_name %in% c("resource_help", "content_help") ~ "content_rec_coal", 
                               var_name %in% value_prop_varname ~ "value_prop", 
                               var_name %in% value_prop_varname_raw ~ "value_prop_raw", 
                               var_name %in% c("languishing", "document_tools", "document_benefits") ~ "value_prop_coal",
                               var_name == "demo_race" ~ "demo_coal", 
                               var_name %in% c("swls_tot", "swls_tot_f", "aniq_tot") ~ "scale", 
                               var_name %in% prolific_demo ~ "prolific_demo", 
                               var_name == "content_recruit" ~ "recruit", 
                               var_name == "duration__in_seconds_" ~ "metadata", 
                               str_detect(var_name, "demo_ethnicity_") ~ "demo_raw", 
                               #str_detect(var_name, "wbpro_") ~ "item", 
                               TRUE ~ item_type))

ltaw_rn_sum %>%
  gt(caption = "Let's Talk About Well-being Standardized Variable Name Summary")
```


# Standardize demographic levels

## re-level based on naming convention
```{r}
# # standardize demographic levels based on naming conventions
# demo_std_l <- std_demo_level(ltaw_df_rn, demo_dic)
# 
# # extract re-leveled dataset
# ltaw_df_rl <- demo_std_l$df_rl
# 
# # extract demographic level summary
# demo_level_sum <- demo_std_l$demo_level_sum
# 
# demo_level_sum %>%
#   gt(caption = "Demographic level summary")
```

# standardize other factors

```{r}
# factor_sum <- factor_summary(ltaw_df_rn,ltaw_rn_sum )
# factor_sum %>% 
#   select(-factor_levels) %>%
#   gt(caption = "A summary of all factor variables")
```

# write outputs
```{r}
current_date <- format(Sys.Date(), "%d%m%y")

#write.csv(ltaw_df_rn, here("outputs", paste("ltaw_df_std_", current_date, ".csv", sep = "")), row.names = F)
#write.csv(ltaw_rn_sum, here("outputs", paste("ltaw_var_summary_", current_date, ".csv", sep = "")), row.names = F)
#write.csv(demo_level_sum, here("outputs", paste("ltaw_demo_summary_", current_date, ".csv", sep = "")), row.names = F)
```

