---
title: "ltab_codebook"
author: "`r Sys.getenv('USER')`"
date: '`r Sys.Date()`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, message = FALSE)
```

```{r libraries}
library(tidyverse)
library(here)
library(psych)
library(janitor)
library(broom)
library(gt)
library(gtsummary)
library(reactable)
```

# prepping

## load datasets
```{r}
# load survey item coding schema
survey_item_std <- read.csv(here("inputs", "standardize_survey_items.csv"))

# load survey dictionary
survey_std <- read.csv(here("inputs", "standardize_survey.csv"))

# load standardized dataset
ltab_df_std <- read.csv(here("outputs", "ltaw_df_std_050923.csv"))

# load variable naming list
survey_varname_std <- read.csv(here("outputs", "ltaw_var_summary_050923.csv"))

# load item dictionary
ltab_data_scored_dict <- read.csv(here("outputs", "LTAW_data_scored_dict.csv"))
```

## preset
```{r}
scale_list <- c("aniq", "perma", "pwb", "sris", "swls", "wbpro")
```

# Let's Talk About Wellbeing Overview

## Study Design
This is a cross-sectional study launched in Prolific with Qualtrics and CameraTag.  

## Dataset Overview

```{r}
# identify number of observations
obs <- ltab_df_std %>%
  nrow()

unique_obs <- ltab_df_std[unique(ltab_df_std$userid),] %>%
  nrow()

# identify study_idx
study_idx <- unique(ltab_df_std$servey_name)
```

The index of the present dataset is `r study_idx`. This dataset includes `r obs` observations from `r unique_obs` participants.It contains Survey Items, Demographics, Moments, Open-ended questions related to content recommendation and Value prop questions. No repeated measures. 

# Data Overview

## Naming Convention

1. survey items are named based on [scale name]_[item index/subscale name]. 
     
2. The suffix "_r" indicates that the item is a reverse-score item. Reverse-score items are questions rephrased in the opposite way as the other items in the scale. Without further specification, these items are recoded by [max scaled number + 1] - [raw score]. These reverse-score items are used to calculate the (sub)scale score instead of the raw scores.   
     
3. The suffix"_combine" indicates that the item is a coalesced item. If participants are asked to select all applicable options for a question, then its corresponding coalesced item will indicates if they selected more than one option or the single option that they selected. 

## Data Screening Flags
```{r}

item_dict <- ltab_data_scored_dict %>%
  filter(is.na(value))


screen_df <- survey_varname_std %>%
  select(-scale_prefix, -item_suffix) %>%
  left_join(select(item_dict, variable, label), by = c("ori_name" = "variable")) %>%
  filter(item_type == "data_screen_flags")

screen_df %>%
    reactable(striped = TRUE,
            filterable = TRUE,   resizable = TRUE)

```

## Analysis Recommendation

1. Use variables included in the subscale overview as metrics of psychological status. 
2. Include cases marked by the flag_rec_include in the analysis


# Scale Level Overview

```{r}

scale_df <- survey_std %>%
  filter(scale_prefix %in% scale_list) %>%
  filter(item_suffix == "") %>%
  select(-item_suffix, -calculation, -range)

scale_df %>%
  reactable(striped = TRUE,
            filterable = TRUE,   resizable = TRUE)

```

# Subscale level overview

```{r}
subscale_name <- survey_varname_std$var_name[survey_varname_std$item_type == "subscale" | survey_varname_std$item_type == "scale"]

subscale_df <- ltab_df_std %>%
  select(all_of(subscale_name)) %>%
  select(-swls_tot_f) %>%
  gather(subscale, value) %>%
  nest(-subscale) %>%
  mutate(des = map(data, ~describe(.x$value))) %>%
  unnest(des) %>%
  select(-data, -vars) %>%
  mutate(missing_perc = (obs - n)/n * 100) %>%
  mutate_if(is.numeric,round, 3) %>%
  select(-trimmed, -mad, -se) %>%
  left_join(select(survey_std, var_name, function.,calculation), by = c("subscale" = "var_name")) %>%
  relocate(function., .before = 2)

subscale_df %>%
  reactable(striped = TRUE,
            filterable = TRUE,   resizable = TRUE)
```

# Item-level summary

```{r}
# subset with std data with item variables
std_item_names <- survey_varname_std$var_name[survey_varname_std$item_type == "item" | survey_varname_std$item_type == "recoded_item"]

item_df <- ltab_df_std %>%
  select(all_of(std_item_names))

# calculate percentage of missing 
scaleitem_missing <- map_dbl(item_df, ~ 100*mean(is.na(.)))

ltab_var_df <- data.frame(question_id = std_item_names) %>%
  mutate(missing_perc = round(scaleitem_missing, 3), 
         rept_measure = FALSE)

# merge with standardized coding schema
ltab_item_sum <- survey_item_std %>%
  mutate(include = case_when(question_id %in% std_item_names ~ TRUE,
                             TRUE ~ FALSE)) %>%
  left_join(ltab_var_df, by = "question_id")

ltab_item_sum %>%
  reactable(striped = TRUE,
            filterable = TRUE,   resizable = TRUE)
```

# demographic summary

## Assessed by survey item

The following demographic categories were assessed by the survey item
```{r}
survey_varname_std$var_name[survey_varname_std$item_type == "demo" | survey_varname_std$item_type == "demo_coal"]
```

Categories within each demographic variable:  
```{r}
ltab_demo_sum <- ltab_df_std %>%
  select(contains("demo_")) %>%
  pivot_longer(cols = everything(), names_to = "demo_var", values_to = "category") %>%
  group_by(demo_var, category) %>% 
  summarise(n = n(), .groups = "drop") %>%
  group_by(demo_var) %>%
  mutate(percentage = round(n / sum(n) * 100, 3)) %>%
  filter(category !="")

ltab_demo_sum %>%
  reactable(striped = TRUE,
            filterable = TRUE,   resizable = TRUE)
```

## Assessed by prolific

```{r}
(prolific_demo_name <- survey_varname_std$var_name[survey_varname_std$item_type == "prolific_demo"])
```

age_prolific is a numeric variable  


Categories within each demographic variable: 
```{r}
ltab_demo_sum <- ltab_df_std %>%
  select(prolific_demo_name) %>%
  select(-age_prolific) %>%
  pivot_longer(cols = everything(), names_to = "demo_var", values_to = "category") %>%
  group_by(demo_var, category) %>% 
  summarise(n = n(), .groups = "drop") %>%
  group_by(demo_var) %>%
  mutate(percentage = round(n / sum(n) * 100, 3)) %>%
  filter(category !="")

ltab_demo_sum %>%
  reactable(striped = TRUE,
            filterable = TRUE,   resizable = TRUE)
```

# all data label

item_type explanation:   
     
* metadata: data related to survey status (e.g. survey id; duration; progress). 
   
* demo: demographic related data. 
   
* demo_raw: options for demographic related questions that allow participants to select applicable options. Each variable indicates if participant selected the option as one of their responses.   
   
* item: psychological survey item. In most cases, they shouldn't be included in any analysis. 
   
* attention_check: filler questions aims to check participants attention. No need to be included in any analysis. 
   
* subscale: scores that should be used as the metrics of a dimension of a psychological construct(status).
   
* scale: scores that should be used as the overall metrics of a psychological construct(status). 
   
* prompt_meta: data related to moment collection.  
    
* content_rec_raw: options for questions related to content recommendation that allow participants to select applicable options. Each variable indicates if participant selected the option as one of their responses.
     
* value_prop_raw: options for questions related to value proposition that allow participants to select applicable options. Each variable indicates if participant selected the option as one of their responses.   
     
* prolific_metadata: data related to survey status that were recorded by the survey platform (Prolific).  
     
* prolific_demo: demographic data that were collected by the survey platform (Prolific).   
      
* data_screen_flags: records of data cleaning & screening process
     
* coalesced responses: coalesced responses for questions that allow participants to select all applicable responses.
     
* reverse_scored_item: Reversed score items based on the coding schema of the standardized scales. In most cases, they shouldn't be included in any analysis. 


```{r}
all_data <- survey_varname_std %>%
  select(-scale_prefix, -item_suffix) %>%
  left_join(select(item_dict, variable, label), by = c("ori_name" = "variable"))

all_data %>%
    reactable(striped = TRUE,
            filterable = TRUE,   resizable = TRUE)

#write.csv(all_data, here("outputs", "ltaw_survey_data_dict.csv"), row.names = F)
```

