---
title: "Exploring Community Trend (WHI Pre-pilot) Study Summary 1: Standardization"
author: "`r Sys.getenv('USER')`"
date: '`r Sys.Date()`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, message = FALSE)
```

Standardize Let’s Talk About Well-being dataset based on the variable name and demographic dictionary. 

# Prepping {.tabset}

## load libraries
```{r libraries}
library(tidyverse)
library(here)
library(psych)
library(janitor)
library(broom)
library(gt)
library(gtsummary)
library(dplyr)
```

## load data
```{r}
# load cleaned datasets
whi_df <- read.csv(here("outputs", "whi_data_scored.csv"))

```

## org_varname function
A function to organize variable names
```{r}
org_varname <- function(dataframe){
  scale_df <- data.frame("var_name" = colnames(dataframe), 
                            "var_class" = sapply(dataframe, class)) %>%
  separate(var_name, into = c("scale_prefix", "item_suffix"), sep = "_", remove = FALSE, extra = "merge") %>%
  mutate(num_item = grepl("\\d", item_suffix))
  
  scale_df <- scale_df %>%
  mutate(item_type = case_when(
    str_detect(scale_prefix, "(?i)Consent") ~ "metadata",
    str_detect(scale_prefix, "(?i)recording") ~ "metadata",
    str_detect(scale_prefix, "(?i)demo") ~ "demo",
    str_detect(scale_prefix, "(?i)flag") ~ "data_screen_flags",
    str_detect(scale_prefix, "(?i)max") ~ "data_screen_flags",
    str_detect(scale_prefix, "(?i)Attention") ~ "attention_check",
    str_detect(scale_prefix, "(?i)Prompt") ~ "prompt",
    str_detect(item_suffix, "(?i)_tot") ~ "subscale",
    str_detect(item_suffix, "(?i)_R") ~ "recoded_item",
    str_detect(item_suffix, "(?i)prolific") ~ "prolific_metadata", 
    str_detect(scale_prefix, "(?i)flag") ~ "data_screen_flags",
    var_name == "Duration__in_seconds_" ~ "metadata",
    num_item ~ "item",
    !is.na(item_suffix) & !num_item & var_class != "character" ~ "subscale",
    scale_prefix == "outro" ~ "feedback", 
    TRUE ~ "metadata"
  ))
  
  return(select(scale_df, -num_item))
}
```

## compare_varname function

A function to compare the variable names from the dataset against the dictionary
```{r}
compare_varname <- function(varname_dic, df){
  
  
  # add row index to the project dataframe
  df_compare <- df %>%
    mutate(row_idx = row_number()) %>%
    select(var_name, var_class, row_idx)
  
  # merge the project dataframe with the variable name dictionary
  matched_df <- varname_dic %>%
    left_join(df_compare, by = c("common_name" = "var_name"), suffix =c("_dic", "_df"))
  
  # identify unmatched rows from the project dataframe
  unmatched_df <- anti_join(df_compare, varname_dic, by = c("var_name" = "common_name"))
  
  return(list(matched_df = matched_df, unmatched_df = unmatched_df))
  
}
```

## rename_match function

A function to rename variable names based on the naming convention

```{r}
rename_var_std <- function(df) {
  df %>%
    #clean_names() %>% # use janitor::clean_names for a more thorough cleanup
    rename_with(~ tolower(gsub(".", "_", .x, fixed = TRUE))) %>% # replace "." by "_" and change everything into lower case
    rename_with(~ str_replace_all(., "(\\d+)", function(x) sprintf("%02d", as.numeric(x))), .cols = everything()) #add padding 0 to all numeric index
    #rename_with(~ str_replace_all(., "\\d+", function(x) sprintf("%02d", as.numeric(x))), .cols = everything())
}
```

## std_var_name function

A function to standardize variable names. It renames matched variables based on the dictionary and renames unmatched variables based on the naming convention
```{r}
std_var_name <- function(df, varname_dic){
  
  # organize variable name
  varname_df <- org_varname(df)
  
  # rename matched variable
  # compare the project variable against the variable name dictionary
  df_compare_results <- compare_varname(varname_dic, varname_df)
  # subset a dataset with comparable variables 
  matched_df <- df_compare_results$matched_df
  # rename the variable names based on the matched names from the dictionary
  colnames(df)[matched_df$row_idx] <- matched_df$std_varname
  
  # rename unmatched variable
  df_rn <- rename_var_std(df)
  
  # organize standardized variable name 
  varname_df_std <- org_varname(df_rn)  %>% # create a summary table
    mutate(ori_name = varname_df$var_name) %>%
    mutate(match_dic = row_number() %in% matched_df$row_idx)
  
  std_results <- list(df_std = df_rn, varname_sum_std = varname_df_std)
  return(std_results)
}
```

## value_between

```{r}
value_between <- function(vec, start, end){
  
  result <- vec[(which(vec == start)):(which(vec == end))]
  
  return(result)
}
```

## as_factor_same

```{r}
as_factor_same <- function(df, var_names){
  df_f <- df %>%
    mutate(across(var_names, ~as.factor(.)))
  
  return(df_f)
}
```

## as_factor_recode
```{r}
as_factor_recode <- function(df, var_names, vec_recode_pattern){
  df_f <- df %>%
    mutate(across(var_names, ~as.factor(.))) %>%
    mutate(across(var_names, ~recode_factor(., !!!vec_recode_pattern)))
  
  return(df_f)
}
```


## demo_summary function

A function to summarize demographic variables and levels
```{r}
demo_summary <- function(dataframe){
  
  # convert all demo variables into factors
  demo_df <- dataframe %>%
    select(contains("Demo")) %>%
    mutate_all(as.factor)
  
  # extract levels and their total number
  demo_levels <- sapply(demo_df, levels)
  demo_levels_n <- sapply(demo_levels, length)
  
  # transform the levels into a dataframe
  demo_levels_df <- map_dfr(names(demo_levels), ~ data.frame(demo_var = .x, levels = unlist(demo_levels[[.x]]))) %>%
    nest(demo_levels = c(levels), .by = demo_var, .key = "demo_level_name")
  
  # combine
  demo_sum <- data.frame("demo_var" = colnames(demo_df), 
                                 "demo_levels_n" = demo_levels_n) %>%
    left_join(demo_levels_df, by = "demo_var") %>%
    unnest(demo_levels)
  
  return (demo_sum)
}
```

## compare_demo function

A function to compare demographic levels against the dictionary
```{r}
compare_demo <- function(demo_dic, df){
  
  # organize demographic levels
  demo_level_df <- demo_summary(df)

  # compare demographic variables and number of levels
  demo_dic_sum <- demo_dic %>%
    nest(levels) %>%
    rename(demo_levels = data)
  
  demo_df_sum <- demo_level_df %>%
    nest(levels) %>%
    rename(demo_levels = data)
  
  demo_sum_compare <- demo_dic_sum %>%
    full_join(demo_df_sum, by = c("common_name" = "demo_var"), suffix = c("_dic", "_df"))
  
  # check demographic variables that are not included in the dataset 
  missing_demo_var <- anti_join(demo_dic_sum, demo_df_sum, by = c("common_name" = "demo_var"))
  
  if(nrow(missing_demo_var) != 0) {
    missing_demo <- sprintf("Demographic variables that are not included in the dataset: %s \n", missing_demo_var$common_name)
    cat(missing_demo)
  } else {
    cat("All demographic variables from the dictionary are included in the dataset \n")
  } 
  
  # check demographic variables that are not included in the dictionary
  additional_demo_var <- anti_join(demo_df_sum, demo_dic_sum, by = c("demo_var" = "common_name"))

  if(nrow(additional_demo_var) != 0) {
    missing_demo <- sprintf("Demographic variables that are not included in the dictionary: %s \n", additional_demo_var$demo_var)
    cat(missing_demo)
  } else {
    cat("No additional demographic varialbe is included in the dataset \n")
  } 
  
  
  # compare items & subscales
  demo_compare_eval <- demo_sum_compare %>% 
  mutate(
         same_demo_n = identical(demo_levels_n_dic, demo_levels_n_df), # check if the index of items are identical
         same_demo_level = identical(demo_levels_dic, demo_levels_df)
         ) # check if the numes of subscales are identical
  
  return(demo_compare_eval)
  
}
```

## std_demo_level
```{r}
std_demo_level <- function(df, demo_dic) {
  
  # standardize demographic levels based on naming conventions
  df_rl <- df %>%
    mutate_at(vars(starts_with("demo_")), ~replace(., . == "", NA)) %>% # replace empty strings with NA
    mutate_at(vars(starts_with("demo_")), ~str_to_lower(.)) # transform all categories into lower_case
  
  # compare demographic levels against dictionary
  demo_sum_compare <- compare_demo(demo_dic, df_rl)
  
  # generate a summary of demographic levels
  demo_level_sum <- demo_summary(df_rl)
  
  std_demo_l <- list(df_rl = df_rl, demo_level_sum = demo_level_sum)
  
  return(std_demo_l)
  
}
```

## factor_summary
```{r}
factor_summary <- function(df, varname_sum){
  
  # extract factor name
  factor_var_df <- varname_sum %>%
  filter(var_class == "factor") %>%
  filter(!str_detect(var_name, "demo_"))

  factor_varname <- factor_var_df$var_name
  
  # extract factor data
  factor_df <- df %>%
  select(one_of(factor_varname))

  # extract levels and their total number
  factor_levels <- sapply(factor_df, levels)
  factor_levels_n <- sapply(factor_levels, length)
  
   factor_levels_df <- map_dfr(names(factor_levels), ~ data.frame(var_name = .x, levels = unlist(factor_levels[[.x]]))) %>%
    nest(factor_levels = c(levels), .by = var_name, .key = "demo_level_name")
  
  # combine
  factor_sum <- data.frame("var_name" = factor_var_df$var_name, 
                                 "factor_levels_n" = factor_levels_n) %>%
    left_join(factor_levels_df, by = "var_name")
  
  return(factor_sum)
  
}

```

# A summery of orignal variable names

```{r}
ori_varname <- org_varname(whi_df)

ori_varname %>%
  gt(caption = "WHI Pre-pilot Original Variables")
```

# standardize prolific data
```{r}
whi_df <- whi_df %>%
  mutate(Age_prolific = case_when(Age_prolific == "DATA_EXPIRED" ~ NA, 
                                  TRUE ~ as.numeric(Age_prolific))) %>%
  dplyr::mutate(across(Sex_prolific:Employment.status_prolific, # recode binary variables into a Boolean
                ~ case_when(
                  . == "DATA_EXPIRED" ~ NA,
                  . == "Prefer not to say" ~ NA,
                  TRUE ~ .
                ))) %>%
  dplyr::mutate(across(c(Status_prolific, Current.u.s.state.of.residence_prolific, Fluent.languages_prolific, Sex_prolific:Employment.status_prolific), # recode binary variables into a Boolean
                ~ as.factor(.)
                ))
```


# change item class

## recode character to factor without changing levels
```{r}
# check all character variables
whi_char_df <- ori_varname %>%
  filter(var_class == "character")

whi_char_df %>%
  gt(caption = "All character variables")

# identify variables that need to be recoded as factors
chr_col_names <- whi_char_df$var_name 

f_varname <- value_between(chr_col_names, "state", "AB_33_rank_3")
f_varname <- f_varname[!grepl("_TEXT$", f_varname)]

# standardize levels
whi_df_rl <- whi_df %>%
  mutate(across(all_of(f_varname), 
                ~ str_to_lower(str_replace_all(., c(" " = "_", "-" = "_")))))

# recode them as factors
whi_df_rc <- as_factor_same(whi_df_rl, f_varname)

```

## recode integer to factor 
```{r}
# check all integer variables
whi_int_df <- ori_varname %>%
  filter(var_class == "integer")

whi_int_df %>%
  gt(caption = "All integer variables")

# recode integer into factor
whi_df_rc <- whi_df_rc %>%
  dplyr::mutate(across(c(QOL_1_prompt, QOL_6_prompt, QOL_8_prompt), 
                ~ case_when(
                  . == 1 ~ "I responded to prompt option 1", 
                  . == 2 ~ "I responded with general detial"
                )),
         QOL_7_prompt == case_when(
           QOL_7_prompt == 1 ~ "I responded to prompt option 1", 
           QOL_7_prompt == 2 ~ "I responded to prompt option 2", 
           QOL_7_prompt == 3 ~ "I responded to prompt option 3", 
           QOL_7_prompt == 4 ~ "I responded with general detail"
         )) %>%
  mutate(across(contains("_prompt"), as.factor))

# check all variable class
org_varname(whi_df_rc) %>%gt()
```

# Standardize variable names

```{r}
# rename variable according to the naming convention
whi_df_rn <- rename_var_std(whi_df_rc) %>%
  rename_with(~ifelse(str_starts(.x, "ab33"), str_replace(.x, "^ab33", "ab_33"), .x)) %>%
  rename(current_us_state_of_residence_prolific = current_u_s_state_of_residence_prolific)

# create a summary table
whi_rn_sum <- org_varname(whi_df_rn) %>%
  mutate(ori_name = names(whi_df_rc))

whi_rn_sum %>% gt()
```

# change item type
```{r}
# extract variables that need to change item type
location_var <- value_between(whi_rn_sum$var_name, "state", "wa_county")
demo_raw_var <- value_between(whi_rn_sum$var_name, "demo_age_house_01", "demo_ethnicity__99")
ab_raw_var <- value_between(whi_rn_sum$var_name, "ab_ak_01", "ab_07_10")
ab_raw_var <- ab_raw_var[!grepl("_TEXT$", f_varname)]
prolific_demo_var <- value_between(whi_rn_sum$var_name, "current_us_state_of_residence_prolific", "employment_status_prolific")
ab_coal_var <- value_between(whi_rn_sum$var_name, "arkansas_hc_location_ab_ak", "ab_33_rank_03")
scale_tot_var <- c("qol_tot", "sdoh_tot", "ch_tot", "prbq_tot")


whi_rn_sum <- whi_rn_sum %>%
  mutate(item_type = case_when(var_name %in% location_var ~ "location", 
                               var_name %in% demo_raw_var ~ "demo_raw", 
                               var_name %in% ab_raw_var ~ "assets_barriers_raw", 
                               var_name %in% prolific_demo_var ~ "prolific_demo", 
                               var_name %in% ab_coal_var ~ "assets_barriers_coal", 
                               var_name %in% scale_tot_var ~ "scale", 
                               var_name == "survey_name" ~ "metadata", 
                               var_name == "demo_race" ~ "demo_coal",
                               TRUE ~ item_type))

whi_rn_sum %>%
  gt(caption = "Let's Talk About Well-being Standardized Variable Name Summary")
```

# write outputs
```{r}
current_date <- format(Sys.Date(), "%d%m%y")

#write.csv(whi_df_rn, here("outputs", paste("whi_df_std_", current_date, ".csv", sep = "")), row.names = F)
#write.csv(whi_rn_sum, here("outputs", paste("whi_var_summary_", current_date, ".csv", sep = "")), row.names = F)
#write.csv(demo_level_sum, here("outputs", paste("ltaw_demo_summary_", current_date, ".csv", sep = "")), row.names = F)
```

