---
title: "prompt_language_moment_standardize"
author: "Bernice Cheung"
date: "2023-09-13"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepping {.tabset}

## load libraries
```{r libraries}
library(tidyverse)
library(here)
library(psych)
library(janitor)
library(broom)
library(gt)
library(gtsummary)
library(dplyr)
```

## load data
```{r}
# load cleaned datasets
moment_df <- read.csv(here("outputs", "prompt_language_moment_clean.csv"))

# load prompt dictionary
prompt_dic <- read.csv(here("outputs", "prompt_moment_dict.csv"))
```

## org_varname function
A function to organize variable names
```{r}
org_moment_varname <- function(dataframe){
  scale_df <- data.frame("var_name" = colnames(dataframe), 
                            "var_class" = sapply(dataframe, class)) %>%
  separate(var_name, into = c("scale_prefix", "item_suffix"), sep = "_", remove = FALSE, extra = "merge") %>%
  mutate(num_item = grepl("\\d", item_suffix))
  
  scale_df <- scale_df %>%
  mutate(item_type = case_when(
    str_detect(scale_prefix, "(?i)flag") ~ "data_screen_flags",
    var_name %in% c("duration_moment", "number_of_words", "words_per_minute") ~ "moment_engagement", 
    var_name == "transcript" ~ "moment",
    str_detect(scale_prefix, "(?i)Prompt") ~ "prompt_meta",
    var_name %in% c("prompt_category", "prompt_construct", "prompt_framing") ~ "prompt_character",
    TRUE ~ "metadata"
  ))
  
  return(select(scale_df, -num_item))
}
```

## rename_match function

A function to rename variable names based on the naming convention

```{r}
rename_var_std <- function(df) {
  df %>%
    #clean_names() %>% # use janitor::clean_names for a more thorough cleanup
    rename_with(~ tolower(gsub(".", "_", .x, fixed = TRUE))) %>% # replace "." by "_" and change everything into lower case
    rename_with(~ str_replace(., "(\\d+)", function(x) sprintf("%02d", as.numeric(x))), .cols = everything()) #add padding 0 to all numeric index
}
```

## as_factor_same
```{r}
as_factor_same <- function(df, var_names){
  df_f <- df %>%
    mutate(across(var_names, ~as.factor(.)))
  
  return(df_f)
}
```

## factor_summary
```{r}
factor_summary <- function(df, varname_factor){
  
  # extract factor data
  factor_df <- df %>%
  select(all_of(varname_factor))

  # extract levels and their total number
  factor_levels <- sapply(factor_df, levels)
  factor_levels_n <- sapply(factor_levels, length)
  
  # organize factor levels into a dataframe with level index
  factor_levels_df <- map_dfr(names(factor_levels), ~ data.frame(var_name = .x, levels = unlist(factor_levels[[.x]]))) %>%
    group_by(var_name) %>%
    mutate(level_idx = row_number())
  
  return(factor_levels_df)
  
}

```

# add prompt_id 
merge prompt id from the prompt dictionary
```{r}
moment_df_merge <- moment_df %>%
  left_join(select(prompt_dic, prompt_idx, prompt_id), by = "prompt_idx")
```

# A summery of orignal variable names

```{r}
ori_varname <- org_moment_varname(moment_df_merge)

ori_varname %>%
  gt(caption = "Prolific Prompt Language Original Variables")

ori_var <- ori_varname$var_name
```

# change item class

```{r}
# transform duration into numeric 
prompt_df_rc <- moment_df_merge %>%
  mutate(duration_moment = ifelse(duration_moment == "none", NA, duration_moment), 
         duration_moment = as.numeric(duration_moment), 
         words_per_minute = ifelse(words_per_minute == "none", NA, words_per_minute), 
         words_per_minute = as.numeric(words_per_minute))

# change moment character variables into factors
f_name<- c("prompt_category", "prompt_construct", "prompt_framing")
# recode them as factors
prompt_df_rc <- as_factor_same(prompt_df_rc, f_name)

# check remaining character variables
org_moment_varname(prompt_df_rc) %>%
  filter(var_class == "character") %>%
  gt()
```

## standardize variable names & prompt_idx
```{r}
# rename variable according to the naming convention
prompt_df_rn <- rename_var_std(prompt_df_rc) %>%
  mutate(prompt_idx = str_replace_all(prompt_idx, "Prompt_", "")) %>%
  mutate(prompt_idx = str_replace_all(prompt_idx, "(?<=^|_)([0-9]{1}(?=[^0-9]|$))", "0\\1")) %>%
  mutate(prompt_idx = tolower(prompt_idx))

prompt_df_rn_sum <- org_moment_varname(prompt_df_rn) %>%
    mutate(ori_var = ori_var) 


prompt_df_rn_sum %>%
  gt()
```

# create a moment_codebook

## summarize factor
```{r}
# organize factors and their levels into a dataframe
factor_levels_df <- factor_summary(prompt_df_rn, f_name)
```

```{r}
# merge variable summary with data dictionary
prompt_std_dic <- prompt_df_rn_sum %>%
  full_join(factor_levels_df, by = "var_name") %>%
  mutate(label = NA,
         data_source = "moment_transcript") %>%
  select(var_name, var_class, item_type, label, levels, level_idx, ori_name = ori_var, data_source, scale_prefix, item_suffix)
```

```{r}
# add some labels
prompt_std_dic$label[which(prompt_std_dic$var_name == "userid")] <- "User IDs that can be used for data merging"
prompt_std_dic$label[which(prompt_std_dic$var_name == "prompt_idx")] <- "The prompt index created specifically for the prompt language study"
prompt_std_dic$label[which(prompt_std_dic$var_name == "prompt_text")] <- "The prompt given to the participants"
prompt_std_dic$label[which(prompt_std_dic$var_name == "transcript")] <- "transcription of the moment data"
prompt_std_dic$label[which(prompt_std_dic$var_name == "flag_invalid_moment")] <- "TRUE if either the duration is none or the item type is PromptTest"
prompt_std_dic$label[which(prompt_std_dic$var_name == "flag_dup_moment")] <- "TRUE if there are more than one identical item type within a userid (all duplicated moments are marked); NA if the moments are invalid"
prompt_std_dic$label[which(prompt_std_dic$var_name == "prompt_id")] <- "The prompt index from the prompt dictionary"
```

```{r}
# write outputs
current_date <- format(Sys.Date(), "%d%m%y")

#write.csv(prompt_df_rn, here("outputs", paste("prolific_prompt_moment_std_", current_date, ".csv", sep = "")), row.names = F)
#write.csv(prompt_std_dic, here("outputs", paste("prolific_prompt_moment_codebook_", current_date, ".csv", sep = "")), row.names = F)
#write.csv(demo_level_sum, here("outputs", paste("ltaw_demo_summary_", current_date, ".csv", sep = "")), row.names = F)

```
