---
title: "prompt_language_standardize"
author: "Bernice Cheung"
date: "2023-09-08"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document:
    toc: true
    toc_depth: 2
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, message = FALSE)
```

Standardize Let’s Talk About Well-being dataset based on the variable name and demographic dictionary. 

# Prepping {.tabset}

## load libraries
```{r libraries}
library(tidyverse)
library(here)
library(psych)
library(janitor)
library(broom)
library(gt)
library(gtsummary)
library(dplyr)
```

## load data
```{r}
# load cleaned datasets
prompt_df <- read.csv(here("outputs", "prolific_prompt_data_scored.csv"))
prompt_df_dic <- read.csv(here("outputs", "prolific_prompt_data_scored_dict.csv"))
# load item variable dictionary
#varname_dic <- read.csv(here("outputs", "varname_dic_030823.csv"))

# load demographic level dictionary
#demo_dic <- read.csv(here("outputs", "demo_level_dic_020823.csv"))
```

## org_varname function
A function to organize variable names
```{r}
org_varname <- function(dataframe){
  scale_df <- data.frame("var_name" = colnames(dataframe), 
                            "var_class" = sapply(dataframe, class)) %>%
  separate(var_name, into = c("scale_prefix", "item_suffix"), sep = "_", remove = FALSE, extra = "merge") %>%
  mutate(num_item = grepl("\\d", item_suffix))
  
  scale_df <- scale_df %>%
  mutate(item_type = case_when(
    str_detect(scale_prefix, "(?i)Consent") ~ "metadata",
    str_detect(scale_prefix, "(?i)recording") ~ "metadata",
    str_detect(scale_prefix, "(?i)demo") ~ "demo",
    str_detect(scale_prefix, "(?i)flag") ~ "data_screen_flags",
    str_detect(scale_prefix, "(?i)max") ~ "data_screen_flags",
    str_detect(scale_prefix, "(?i)Attention") ~ "attention_check",
    str_detect(scale_prefix, "(?i)Prompt") ~ "prompt_meta",
    str_detect(item_suffix, "(?i)Prompt") ~ "prompt_meta",
    str_detect(item_suffix, "(?i)combine") ~ "coalesced responses",
    str_detect(item_suffix, "(?i)Text") ~ "supplement responses",
    str_detect(item_suffix, "(?i)prolific") ~ "prolific_metadata",
    str_detect(item_suffix, "(?i)_tot") ~ "subscale",
    str_detect(item_suffix, "(?i)_R$") ~ "reverse_scored_item",
    str_detect(scale_prefix, "(?i)flag") ~ "data_screen_flags",
    num_item ~ "item",
    !is.na(item_suffix) & !num_item & var_class != "character" ~ "subscale",
    scale_prefix == "outro" ~ "feedback", 
    TRUE ~ "metadata"
  ))
  
  return(select(scale_df, -num_item))
}
```

## compare_varname function

A function to compare the variable names from the dataset against the dictionary
```{r}
compare_varname <- function(varname_dic, df){
  
  
  # add row index to the project dataframe
  df_compare <- df %>%
    mutate(row_idx = row_number()) %>%
    select(var_name, var_class, row_idx)
  
  # merge the project dataframe with the variable name dictionary
  matched_df <- varname_dic %>%
    left_join(df_compare, by = c("common_name" = "var_name"), suffix =c("_dic", "_df"))
  
  # identify unmatched rows from the project dataframe
  unmatched_df <- anti_join(df_compare, varname_dic, by = c("var_name" = "common_name"))
  
  return(list(matched_df = matched_df, unmatched_df = unmatched_df))
  
}
```

## rename_match function

A function to rename variable names based on the naming convention

```{r}
rename_var_std <- function(df) {
  df %>%
    #clean_names() %>% # use janitor::clean_names for a more thorough cleanup
    rename_with(~ tolower(gsub(".", "_", .x, fixed = TRUE))) %>% # replace "." by "_" and change everything into lower case
    rename_with(~ str_replace_all(., "(\\d+)", function(x) sprintf("%02d", as.numeric(x))), .cols = everything()) #add padding 0 to all numeric index
}
```

## std_var_name function

A function to standardize variable names. It renames matched variables based on the dictionary and renames unmatched variables based on the naming convention
```{r}
std_var_name <- function(df, varname_dic){
  
  # organize variable name
  varname_df <- org_varname(df)
  
  # rename matched variable
  # compare the project variable against the variable name dictionary
  df_compare_results <- compare_varname(varname_dic, varname_df)
  # subset a dataset with comparable variables 
  matched_df <- df_compare_results$matched_df
  # rename the variable names based on the matched names from the dictionary
  colnames(df)[matched_df$row_idx] <- matched_df$std_varname
  
  # rename unmatched variable
  df_rn <- rename_var_std(df)
  
  # organize standardized variable name 
  varname_df_std <- org_varname(df_rn)  %>% # create a summary table
    mutate(ori_name = varname_df$var_name) %>%
    mutate(match_dic = row_number() %in% matched_df$row_idx)
  
  std_results <- list(df_std = df_rn, varname_sum_std = varname_df_std)
  return(std_results)
}
```

## value_between

```{r}
value_between <- function(vec, start, end){
  
  result <- vec[(which(vec == start)):(which(vec == end))]
  
  return(result)
}
```

## as_factor_same

```{r}
as_factor_same <- function(df, var_names){
  df_f <- df %>%
    mutate(across(var_names, ~as.factor(.)))
  
  return(df_f)
}
```

## as_factor_recode
```{r}
as_factor_recode <- function(df, var_names, vec_recode_pattern){
  df_f <- df %>%
    mutate(across(var_names, ~as.factor(.))) %>%
    mutate(across(var_names, ~recode_factor(., !!!vec_recode_pattern)))
  
  return(df_f)
}
```


## demo_summary function

A function to summarize demographic variables and levels
```{r}
demo_summary <- function(dataframe){
  
  # convert all demo variables into factors
  demo_df <- dataframe %>%
    select(contains("Demo")) %>%
    mutate_all(as.factor)
  
  # extract levels and their total number
  demo_levels <- sapply(demo_df, levels)
  demo_levels_n <- sapply(demo_levels, length)
  
  # transform the levels into a dataframe
  demo_levels_df <- map_dfr(names(demo_levels), ~ data.frame(demo_var = .x, levels = unlist(demo_levels[[.x]]))) %>%
    nest(demo_levels = c(levels), .by = demo_var, .key = "demo_level_name")
  
  # combine
  demo_sum <- data.frame("demo_var" = colnames(demo_df), 
                                 "demo_levels_n" = demo_levels_n) %>%
    left_join(demo_levels_df, by = "demo_var") %>%
    unnest(demo_levels)
  
  return (demo_sum)
}
```

## compare_demo function

A function to compare demographic levels against the dictionary
```{r}
compare_demo <- function(demo_dic, df){
  
  # organize demographic levels
  demo_level_df <- demo_summary(df)

  # compare demographic variables and number of levels
  demo_dic_sum <- demo_dic %>%
    nest(levels) %>%
    rename(demo_levels = data)
  
  demo_df_sum <- demo_level_df %>%
    nest(levels) %>%
    rename(demo_levels = data)
  
  demo_sum_compare <- demo_dic_sum %>%
    full_join(demo_df_sum, by = c("common_name" = "demo_var"), suffix = c("_dic", "_df"))
  
  # check demographic variables that are not included in the dataset 
  missing_demo_var <- anti_join(demo_dic_sum, demo_df_sum, by = c("common_name" = "demo_var"))
  
  if(nrow(missing_demo_var) != 0) {
    missing_demo <- sprintf("Demographic variables that are not included in the dataset: %s \n", missing_demo_var$common_name)
    cat(missing_demo)
  } else {
    cat("All demographic variables from the dictionary are included in the dataset \n")
  } 
  
  # check demographic variables that are not included in the dictionary
  additional_demo_var <- anti_join(demo_df_sum, demo_dic_sum, by = c("demo_var" = "common_name"))

  if(nrow(additional_demo_var) != 0) {
    missing_demo <- sprintf("Demographic variables that are not included in the dictionary: %s \n", additional_demo_var$demo_var)
    cat(missing_demo)
  } else {
    cat("No additional demographic varialbe is included in the dataset \n")
  } 
  
  
  # compare items & subscales
  demo_compare_eval <- demo_sum_compare %>% 
  mutate(
         same_demo_n = identical(demo_levels_n_dic, demo_levels_n_df), # check if the index of items are identical
         same_demo_level = identical(demo_levels_dic, demo_levels_df)
         ) # check if the numes of subscales are identical
  
  return(demo_compare_eval)
  
}
```

## std_demo_level
```{r}
std_demo_level <- function(df, demo_dic) {
  
  # standardize demographic levels based on naming conventions
  df_rl <- df %>%
    mutate_at(vars(starts_with("demo_")), ~replace(., . == "", NA)) %>% # replace empty strings with NA
    mutate_at(vars(starts_with("demo_")), ~str_to_lower(.)) # transform all categories into lower_case
  
  # compare demographic levels against dictionary
  demo_sum_compare <- compare_demo(demo_dic, df_rl)
  
  # generate a summary of demographic levels
  demo_level_sum <- demo_summary(df_rl)
  
  std_demo_l <- list(df_rl = df_rl, demo_level_sum = demo_level_sum)
  
  return(std_demo_l)
  
}
```

## factor_summary
```{r}
factor_summary <- function(df, varname_factor){
  
  # extract factor data
  factor_df <- df %>%
  select(all_of(varname_factor))

  # extract levels and their total number
  factor_levels <- sapply(factor_df, levels)
  factor_levels_n <- sapply(factor_levels, length)
  
  # organize factor levels into a dataframe with level index
  factor_levels_df <- map_dfr(names(factor_levels), ~ data.frame(var_name = .x, levels = unlist(factor_levels[[.x]]))) %>%
    group_by(var_name) %>%
    mutate(level_idx = row_number())
  
  return(factor_levels_df)
  
}

```


# A summery of orignal variable names

```{r}
ori_varname <- org_varname(prompt_df)

ori_varname %>%
  gt(caption = "Prolific Prompt Language Original Variables")
```


# change item class

## recode character to factor without changing levels
```{r}
# check all character variables
prompt_char_df <- ori_varname %>%
  filter(var_class == "character")

prompt_char_df %>%
  gt(caption = "All character variables")

# identify variables that need to be recoded as factors
chr_col_names <- prompt_char_df$var_name 

f_varname <- c(value_between(chr_col_names, "Demo_gender", "VIS_Strengths"),
               value_between(chr_col_names, "Goal_1_why", "ValueProp_3_9"),
               value_between(chr_col_names, "ValueProp_4", "ValueProp_5_9"), 
               value_between(chr_col_names, "ValueProp_6_1", "ValueProp_6_10"), 
               value_between(chr_col_names, "ValueProp_7", "ValueProp_8_3"), 
               value_between(chr_col_names, "ValueProp_10_1", "ValueProp_10_10"), 
               value_between(chr_col_names, "Sex_prolific", "Employment.status_prolific"),
               value_between(chr_col_names, "status_prolific_api", "SWLS_tot_f"))

# recode them as factors
prompt_df_rc <- as_factor_same(prompt_df, f_varname)

# check remaining character variables
org_varname(prompt_df_rc) %>%
  filter(var_class == "character") %>%
  gt()
```


## recode integer to factor 
```{r}
# check all integer variables
prompt_int_df <- ori_varname %>%
  filter(var_class == "integer")

prompt_int_df %>%
  gt(caption = "All integer variables")

# recode variables where 2 = no instead of 0 = no
recode_no_var <- c("Prompt_4_1A_confirm", "Prompt_4_2A_confirm", "Prompt_1_2B_confirm", "Prompt_2_2B_confirm", "Prompt_1_3A_confirm", "Prompt_7_1_confirm", "Prompt_7_3_confirm", "Prompt_8_3_confirm", "Prompt_8_4_confirm")

prompt_df_rc_corr <- prompt_df_rc %>%
  mutate(across(all_of(recode_no_var), ~ case_when(. == 2 ~ 0,
                                                   TRUE ~ .)))

# recode integer into boolean
recode_varname_bool <- c(grep("_confirm", colnames(prompt_df_rc_corr), value = TRUE), "Recording_test")

prompt_df_rc_bool <- prompt_df_rc_corr %>%
  mutate(across(all_of(recode_varname_bool), ~ case_when(. == 0 ~ FALSE,
                                                    . == 1 ~ TRUE)))

# check all variable class
org_varname(prompt_df_rc_bool) %>%
  filter(var_class == "integer") %>%
  gt()
```

## change booleans to character

Empty character variables will automatically transformed into a logic. 
```{r}
# check all boolean variables
prompt_b_df <- org_varname(prompt_df_rc_bool) %>%
  filter(var_class == "logical")

prompt_b_varname <- prompt_b_df$var_name
  
prompt_b_df %>%
  gt(caption = "All boolean variables")

# extract variable names that need to be change into other class
empty_var <- prompt_b_varname[str_detect(prompt_b_varname, "thoughts")]

prompt_df_rc_chr <- prompt_df_rc_bool %>%
  mutate(across(all_of(empty_var), as.character))
```



# Standardize variable names

```{r}
# rename variable according to the naming convention
prompt_df_rn <- rename_var_std(prompt_df_rc_chr)

# create a summary table
prompt_rn_sum <- org_varname(prompt_df_rn) %>%
  mutate(ori_name = names(prompt_df_rc_chr))

prompt_rn_sum %>% gt()

# chanage the class for recordeddate
prompt_rn_sum$var_class[which(prompt_rn_sum$var_name == "recordeddate")] <- "date_time"
```

# change item type
```{r}
#metadata variables 
meta_var <-c("survey_part", "duration__in_seconds_")

# value prop option variables
vp_option_var <- c(sprintf("valueprop_03_%02d", seq(1, 9)), sprintf("valueprop_05_%02d", seq(1, 9)), sprintf("valueprop_06_%02d", seq(1, 10)), sprintf("valueprop_08_%02d", seq(1, 3)), sprintf("valueprop_10_%02d", seq(1, 10)))

# adjust item type
prompt_rn_sum_new <- prompt_rn_sum %>%
  mutate(item_type = case_when(
    var_name %in% meta_var ~ "metadata", 
    str_detect(var_name, "demo_ethnicity_") ~ "demo_race_options", 
    var_name %in% c("pre_valence_01", "pre_arousal_01") ~ "pre_affect", 
    var_name %in% c("post_valence_01", "post_arousal_01") ~ "post_affect", 
    var_name %in% vp_option_var ~ "value_prop_options",
    var_name == "vis_strengths" ~ "strength_item", 
    var_name == "study_feedback" ~ "feedback",
    var_name == "prolific_nowebsite" ~ "prolific_metadata", 
    var_name == "attempt_order" ~ "metadata", 
    var_name == "sfi_goal_orientation_tot" ~ "subscale", 
    str_detect(var_name, "goal") ~ "goal_item", 
    str_detect(var_name, "valueprop") ~ "value_prop", 
    str_detect(var_name, "topic") ~ "topic_label", 
    str_detect(var_name, "_easy") ~ "prompt_ratings", 
    str_detect(var_name, "_value") ~ "prompt_ratings", 
    str_detect(var_name, "_feelgood") ~ "prompt_ratings", 
    str_detect(var_name, "_feedback") ~ "prompt_feedback", 
    str_detect(var_name, "_thought") ~ "prompt_thoughts", 
    var_name %in% c("swls_tot", "swls_tot_f", "rrs_tot", "sfi_tot") ~ "scale", 
    TRUE ~ item_type
  )) 

prompt_rn_sum_new %>%
  gt(caption = "Prolific Prompt Study Standardized Variable Name Summary")
```

# create a dictionary for all factors

```{r}
# extract factor variables
varname_df_f <- prompt_rn_sum_new %>%
  filter(var_class %in% c("factor"))

varname_f <- varname_df_f$var_name

# organize factors and their levels into a dataframe
factor_levels_df <- factor_summary(prompt_df_rn, varname_f)
```

```{r}
# merge variable summary with data dictionary
prompt_std_dic <- prompt_df_dic %>%
  full_join(prompt_rn_sum_new, by = c("variable" = "ori_name")) %>%
  full_join(factor_levels_df, by = "var_name") %>%
  filter(is.na(value) | (var_class == "integer" & !is.na(value))) %>%
  select(var_name, var_class, item_type, label, value, levels, level_idx, ori_name = variable, scale_prefix, item_suffix)
```

## add session labels
```{r}
# load cleaned dataset for each session
survey1_tidy <- read.csv(here("outputs", "survey1_tidy.csv"))
survey2_tidy <- read.csv(here("outputs", "survey2_tidy.csv"))

# extract varialbe names
survey1_varname <- colnames(survey1_tidy)
survey2_varname <- colnames(survey2_tidy)

# variables unique to survey 1
survey1_unique_var <- setdiff(survey1_varname, survey2_varname)
survey2_unique_var <- setdiff(survey2_varname, survey1_varname)
shared_var <- intersect(survey1_varname, survey2_varname)

prompt_std_dic <- prompt_std_dic %>%
  mutate(survey_part = case_when(ori_name %in% survey1_unique_var ~ "part_01", 
                                 ori_name %in% survey2_unique_var ~ "part_02", 
                                 ori_name %in% shared_var ~ "both",
                                 TRUE ~ NA)) %>%
  relocate(survey_part, .before = ori_name)
```


```{r}
# write outputs
current_date <- format(Sys.Date(), "%d%m%y")

write.csv(prompt_df_rn, here("outputs", paste("prolific_prompt_df_std_", current_date, ".csv", sep = "")), row.names = F)
write.csv(prompt_std_dic, here("outputs", paste("prolific_prompt_codebook_", current_date, ".csv", sep = "")), row.names = F)
#write.csv(demo_level_sum, here("outputs", paste("ltaw_demo_summary_", current_date, ".csv", sep = "")), row.names = F)

```

